\documentclass[a4paper,12pt]{report}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{3}
\input{/usr/share/LaTeX-ToolKit/template.tex}
\begin{document}
\title{Trigonometry}
\author{沈威宇}
\date{\temtoday}
\titletocdoc
\ch{Trigonometry (三角學)}

\ssc{Plane angle (平面角) or angle}
\sssc{Degree, deg, degree of arc, arc degree, or arcdegree (角度)}
One degree, deg, degree of arc, arc degree, or arcdegree, denoted as $^\circ$, is defined such that one full revolution is 360 degrees.

One degree is divided into 60 minutes, minutes of arc, arc minutes, or arcminutes, denoted as a single prime $'$, and one minute is divided into 60 seconds, seconds of arc, arc seconds, or arcseconds, denoted as a double prime $"$.
\sssc{Radian or rad (弧度, 弳, 弳度)}
One radian or rad, denoted as rad or without suffix, is defined as the angle subtended at the center of a unit circle by an arc of unit length, that is $\frac{180}{\pi}^\circ$.
\sssc{Gradian, grad, or gon}
One gradian, grad, or gon, denoted as grad, is defined as $0.9^\circ$.
\sssc{Milliradian, mrad, or mil}
A milliradian, mrad, or mil, denoted as mrad, is defined as $\frac{1}{1000}$ radian.
\sssc{Turn, tr, pla, revolution, rev, cycle, or cyc}
One turn, tr, pla, revolution, rev, cycle, or cyc is defined as $360^\circ$.
\sssc{Generalized angle (廣義角) or directed angle (有向角)}
A generalized angle or directed angle is an angle of any real number with positive value denotes counterclockwise angle and negative value denotes clockwise angle.

Two directed angles are coterminal angles (同界角) if their difference is $2k\pi$ radians with $k\in\bbR$.
\sssc{三角測量}
\begin{itemize}
\item \tb{仰角（angle of elevation）}：仰視目標時，視線與水平線的夾角。
\item \tb{俯角（angle of depression）}：俯視目標時，視線與水平線的夾角。
\item \tb{高度角（altitude angle）}：仰視目標時，即仰角；平視目標時，為零；俯視目標時，即負一乘以俯角。
\item \tb{天頂角（zenith angle）/傾（斜）角（inclination）}：90$^{\circ}$減去高度角。
\item \tb{方位角（azimuth or azimuthal angle）}（地理）：以正北為0$^\circ$，順時針為正。
\item \tb{象限角（reduced bearing）}（地理）：以東南西北某一方位（通常為正北或正南） 為基準，加上向相鄰方位轉向的度數與該相鄰方位，如北35$^\circ$西 代表方位角325$^\circ$、南30$^\circ$西代表方位角210$^\circ$。
\end{itemize}
\ssc{Solid angle (立體角)}
\sssc{Steradian, square radian, or sr (球面度 or 立弳)}
One steradian, square radian, or sr, denoted as sr, is defined as the solid angle subtended at the centre of a unit sphere by a unit area (of any shape) on its surface.
