\documentclass[a4paper,12pt]{article}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{3}
\input{/usr/share/latex-toolkit/template.tex}
\begin{document}
\title{Set Theory}
\author{沈威宇}
\date{\temtoday}
\titletocdoc
\section{Set Theory (集合論)}
\ssc{Basic Notations}
\sssc{Set (集合/集)}
A set, sometimes called a collection or family, is a collection of different things. Theses things are called elements (元素) or members (成員) if the set.

If $a$ is an element of a set $S$, we say that $a$ belongs to $S$ or is in $S$, denoted as $a\in S$; if $b$ is not an element of a set $S$, we say that $b$ does not belong to $S$ or is not in $S$, denoted as $b\notin S$.
\sssc{Specifying a set}
A set may be specified either by listing its elements or by a property that characterizes its elements.

\tb{Roster notation or enumeration notation} specifies a set by listing its elements between braces $\{\,\}$, separated by commas $,$, e.g. $\{\,\}$, $\{x\}$, and $\{x,y\}$. When there is a clear pattern for generating all set elements, one can use ellipses $\ldots$ for abbreviating the notation.

\tb{Set-builder notation} specifies a set as being the set of all elements that satisfy some logical formula. More precisely, if $P(x)$ is a logical formula depending on a variable ⁠$x$, which evaluates to true or false depending on the value of $x$, then
\[\{x\mid P(x)\}\]
or
\[\{x\colon P(x)\}\]
denotes the set of all $x$ of which $P(x)$ is true, in which the vertical bar $\mid$ or the colon $\colon$ is read as "such that", and the the whole formula can be read as "the set of all $x$ such that $P(x)$ is true".
\sssc{Empty set (空集合) and singleton (單元素集合)}
A set with no element in it is called the empty set, denoted as $\varnothing$. A set with one element in it is called a singleton.
\subsubsection{Subset (子集)}
\[A\subseteq B\iff (x\in A\implies x\in B)\]
\sssc{Superset (父集/母集/超集)}
\[B\supseteq A\iff (x\in A\implies x\in B)\]
\subsubsection{Intersection (交集)}
\[A\cap B\coloneq\{x\mid x\in A \land x\in B\}\]
\[\bigcap_{i=1}^n A_i\coloneq\left\{x\middle | \bigwedge_{i=1}^n x\in A_i\right\}\]
\subsubsection{Union (聯集)}
\[A\cup B\coloneq\{x\mid x\in A \lor x\in B\}\]
\[\bigcup_{i=1}^n A_i\coloneq\left\{x\middle | \bigvee_{i=1}^n x\in A_i\right\}\]
\subsubsection{Complement (差集)}
\[A\setminus B\coloneq\{x\mid x\in A \land x\notin B\}\]
\subsubsection{Cartesian product (笛卡爾積)}
\[\prod_{i=1}^nA_i=\{(a_1,a_2,\ldots a_n)\mid a_i\in A_i\},\]
and its elements are called ordered $n$-tuple, in which a $2$-tuple is also called pair and $3$-tuple is also called a triple.
\subsubsection{Power set (冪集合)}
\[2^A\coloneq\{B\mid B\subseteq A\}\]
\sssc{Set scalar arithmetic operation}
\begin{itemize}
\item If $\forall a\in A$, $sa$ is defined, $sA\coloneq\{sa:\,a\in A\}$.
\item If $\forall a\in A$, $a+v$ is defined, $A+v\coloneq\{a+v:\,a\in A\}$.
\end{itemize}
\subsubsection{Kernel (核) of a set of sets}
The kernel of a set $\mathcal {B}\neq \varnothing$ of sets is defined to be:
\[\ker(\mathcal {B})\coloneq\bigcap_{B\in\mathcal {B}}B.\]
\subsection{Zermelo–Fraenkel Set Theory (策梅洛-弗蘭克爾集合論)}
Set theory, or more specifically, Zermelo–Fraenkel set theory, has been the standard way to provide rigorous foundations for all branches of mathematics since the first half of the 20th century. Zermelo–Fraenkel set theory with the axiom of choice included is abbreviated ZFC and denoted $\vdash_{ZFC}$; Zermelo–Fraenkel set theory with the axiom of choice excluded is abbreviated ZF and denoted $\vdash_{ZF}$; sometimes, either of them is denoted $\vdash$.
\subsubsection{Axiom of extensionality (外延公理)}
\[\forall x\forall y[\forall z(z\in x\Leftrightarrow z\in y)\Rightarrow x=y]\]
\subsubsection{Axiom of regularity (正則公理)}
\[\forall x\,(x\neq \varnothing \Rightarrow \exists y(y\in x\land y\cap x=\varnothing ))\]
\subsubsection{Axiom of separation (分類公理)/Axiom schema of specification (規範公理模式)}
Let $\varphi$ be any formula in the language of ZFC with all free variables among $x,z,w_{1},\ldots ,w_{n}$ so that $y$ is not free in 
$\varphi$. Then:
\[\forall z\forall w_{1}\forall w_{2}\ldots \forall w_{n}\exists y\forall x[x\in y\Leftrightarrow ((x\in z)\land \varphi (x,w_{1},w_{2},...,w_{n},z))]\]
This axiom can be used to prove the existence of the empty set, denoted as $\emptyset$ or $\varnothing$.

\tb{Axiom of empty set (空集公理)}：
\[\exists x , \forall y , (y \notin x) \]
The $\varnothing$ is defined as the $x$ above.
\subsubsection{Pairing Axiom (配對公理)}
\[ \forall x\forall y\exists z((x\in z)\land (y\in z))\]
\subsubsection{Union Axiom (聯集公理)}
\[ \forall {\mathcal {F}}\,\exists A\,\forall Y\,\forall x[(x\in Y\land Y\in {\mathcal {F}})\Rightarrow x\in A]\]
Although this formula doesn't directly assert the existence of $\cup \mathcal {F}$, the set $\cup \mathcal {F}$ can be constructed from $A$ in the above using the axiom schema of specification:
\[\cup \mathcal {F}=\{x\in A\mid\exists Y(x\in Y\land Y\in {\mathcal {F}})\}\]
\subsubsection{Axiom schema of replacement (替代公理模式)}
Let $\varphi$ be any formula in the language of ZFC with all free variables among $x,z,w_{1},\ldots ,w_{n}$ so that $B$ is not free in 
$\varphi$. Then:
\[ \forall A\forall w_{1}\forall w_{2}\ldots \forall w_{n}{\bigl [}\forall x(x\in A\Rightarrow \exists !y\,\varphi )\Rightarrow \exists B\ \forall x{\bigl (}x\in A\Rightarrow \exists y(y\in B\land \varphi ){\bigr )}{\bigr ]}\]
\subsubsection{Axiom of Infinity (無窮公理)}
\[ \exists X\left[\exists e(\forall z\,\neg (z\in e)\land e\in X)\land \forall y(y\in X\Rightarrow S(y)\in X)\right]\]
\subsubsection{Power Set Axiom (冪集公理)}
\[\forall A \exists P(A) \forall x (x \in P(A) \leftrightarrow x \subseteq A)\]
The $P(A)$ above is called power set (冪集) and denoted as $2^A$.
\subsubsection{Axiom of Choice (選擇公理)/Axiom of Well-ordering (良序公理)}
\[ \forall X\left[\varnothing \notin X\implies \exists f\colon X\rightarrow \bigcup _{A\in X}A\quad \forall A\in X\,(f(A)\in A)\right]\]
\ssc{Associativity, commutativity, and distributivity of intersection and union}
\begin{itemize}
\item Intersection has commutativity.
\item Union has commutativity.
\item Intersection has associativity and distributivity over intersection.
\item Union has associativity and distributivity over union.
\item Intersection has distributivity over union.
\item Union has distributivity over intersection.
\item Note: $(A\cap B)\cup C$ is not necessarily equal to $A\cap (B\cup C)$
\end{itemize}
\subsection{Russell's paradox or Russell's antinomy (羅素悖論)}
According to the unrestricted comprehension principle, for any sufficiently well-defined property, there is the set of all and only the objects that have that property. Russell's paradox states that, let
\[R\coloneq\{x\mid x\not \in x\}\]
then
\[R\in R\iff R\not \in R.\]

\textit{Statement.} $\vdash_{ZFC}\,\nexists U = \{z \mid z\text{\ is a set\ }\}$.
\begin{proof}\mbox{}\\
Assume that $\exists U = \{z \mid z\text{ is a set}\}$. Let $A=\{x\in U \mid x\notin x\}$. $A \in A\iff (A\in U \land A\notin A)$, but $A\in A \iff \neg A \notin A$, so $A\notin A$, so $\neg(A \in U \land A \notin A)$, so $A \notin U$. $A$ is a set, so $A\in U$. $\Rightarrow\Leftarrow$.
\end{proof}
\ssc{De Morgan's Law (笛摩根定律)}
\sssc{Universe (宇集)}
The universe is now often defined as "when the sets under discussion are all subsets of a given set, the given set is called the universe", denoted as $U$.
\sssc{Complement (補集/餘集)}
If a set $A$ is a subset of a given universe $U$, then the complement of $A$ given $U$ is defined as $U\setminus A$, denoted as $A'$, $\overline{A}$, or $A^C$.
\sssc{Complement laws}
\[(A')'=A\]
\[A\setminus B = A\cap B'\]
\[A\subseteq B \iff B' \subseteq A'\]
\sssc{De Morgan's Law (笛摩根定律)}
\[(A\cup B)'=A'\cap B'\]
\[(A\cap B)'=A'\cup B'\]
\ssc{Binary relation (二元關係) or correspondence}
A binary relation or correspondence $R$ over sets $X$ and $Y$ is a subset of $X\times Y$. The set $X$ is called the domain or set of departure of $R$, and the set $Y$ the codomain or set of destination of $R$. In order to specify the choices of the sets $X$ and $Y$, some authors define a binary relation or correspondence as an ordered triple $(X,Y,G)$, where $G$ is a subset of $X\times Y$ called the graph of the binary relation. The statement $(x,y)\in R$ reads "$x$ is $R$-related to $y$" and is denoted by $xRy$. The domain of definition or active domain of $R$ is the set of all $x\in X$ such that there exists $y\in Y$ such that $xRy$. The codomain of definition, active codomain, image, or range of $R$ is the set of all $y\in Y$ such that there exists $x\in X$ such that $xRy$. The field of $R$ is the union of its domain of definition and its codomain of definition.
\ssc{Equivalence relation (等價關係)}
\sssc{Equicalence relation}
An equivalence relation on a set $X$ is a binary relation $\cong$ on $X$ satisfying the three properties:
\bit
\item Reflexivity:
\[\forall a\in X\colon a\cong a,\]
\item Symmetry:
\[\forall a,b\in X\colon a\cong b\implies b\cong a,\]
\item Transitivity:
\[\forall a,b,c\in X\colon a\cong b\land b\cong a\implies a\cong c.\]
\eit
\sssc{Equivalence classs (等價類)}
An equivalence class of an element $a$ in a set $X$ under an equivalance relation $\cong$ on $X$ is defined as
\[[a]=\{x\in X\mid a\cong x\}.}\]
\sssc{Quotient set (商集)}
The quotient set $Y=X/\cong$ is the set of equivalence classes of elements of $X$ under equivalence relation $\cong$. 
\ssc{Number System (數系)}
\sssc{Natural numbers (自然數)}
The natural numbers $\mathbb{N}$ are defined as positive integers, that is, 1, 2, 3, and so on, denoted as $\mathbb{N}^*$, $\mathbb{N}_1$, or $\mathbb{N}^+$, or non-negative integers, that is, 0, 1, 2, 3, and so on, denoted as $\mathbb{N}^0$ or $\mathbb{N}_0$.

The \tb{Peano axioms (皮亞諾公理), Dedekind–Peano axioms, or Peano postulates (皮亞諾公設)} provide a formal definition of the natural numbers $\mathbb{N}$. Below, we define the natural numbers to include zero; the definition of natural numbers to exclude zero can be constructed by simply replace $0$ with $1$ in the following axioms.
\bit
\item $0$ is a natural number.
\eit
The next four axioms describe the equality relation. Since they are logically valid in first-order logic with equality, they are not considered to be part of "the Peano axioms" in modern treatments.
\bit
\item For every natural number $x$, $x=x$. That is, equality is reflexive.
\item For all natural numbers $x$ and $y$, if $x=y$, then $y=x$. That is, equality is symmetric.
\item For all natural numbers $x$, $y$, and $z$, if $x=y$ and $y=z$, then $x=z$. That is, equality is transitive.
\item For all $a$ and $b$, if $b$ is a natural number and $a=b$, then $a$ is also a natural number. That is, the natural numbers are closed under equality.
\eit
The remaining axioms define the arithmetical properties of the natural numbers. The naturals are assumed to be closed under a single-valued "successor" function $S$.
\bit
\item For every natural number $n$, $S(n)$ is a natural number. That is, the natural numbers are closed under $S$.
\item For all natural numbers $m$ and $n$, if $S(m)=S(n)$, then $m=n$. That is, $S$ is an injection.
\item For every natural number $n$, $S(n)=0$ is false. That is, there is no natural number whose successor is $0$.
\eit
Axioms $1$, $6$, $7$, $8$ define a unary representation of the intuitive notion of natural numbers: the number $1$ can be defined as $S(0)$, $2$ as $S(S(0))$, etc. However, considering the notion of natural numbers as being defined by these axioms, axioms $1$, $6$, $7$, $8$ do not imply that the successor function generates all the natural numbers different from $0$. The intuitive notion that each natural number can be obtained by applying successor sufficiently many times to zero requires an additional axiom, which is sometimes called the \tb{axiom of induction}:
\bit
\item If $K$ is a set such that:
\bit
\item[] $0$ is in $K$, and
\item[] for every natural number $n$, $n$ being in $K$ implies that $S(n)$ is in $K$,
\eit
then $K$ contains every natural number.
\eit

The axiomatization of arithmetic provided by Peano axioms is commonly called \tb{Peano arithmetic (皮亞諾算術)}.
\bit
\item \tb{Addition} is a function that maps two natural numbers to another one. It is defined recursively as:
\[a+0=a,\]
\[a+S(b)=S(a+b).\]
\item \tb{Multiplication} is a function mapping two natural numbers to another one. Given addition, it is defined recursively as:
\[a\cdot 0=0,\]
\[a\cdot S(b)=a+(a\cdot b).\]
\item \tb{Total order}: The field $\mathbb{N}$ is ordered, meaning that there is a total order $\leq$ such that for all natural numbers $x$, $y$ and $z$:
\bit
\item Preservation of order under addition:
\[\forall x,y,z\in\mathbb{N}\colon x\leq y\implies x+z\leq y+z.\]
\item Preservation of order under multiplication:
\[\forall x,y\in\mathbb{N}\colon 0\leq x\land 0\leq y\implies 0\leq xy.\]
\eit
\eit
\sssc{Integers (整數)}
The set of integers $\mathbb{Z}$ can be constructive as follows. First construct the set of natural numbers excluding $0$ according to the Peano axioms, call this $P$. Then construct a set $P^-$ which is disjoint from $P$ and in one-to-one correspondence with $P$ via a function $\psi$. Finally let $0$ be some object not in $P$ or $P^-$. Then the integers are defined to be the union $P\cup P^{-}\cup \{0\}$.
\sssc{Rational numbers (有理數)}
The set of rational numbers (aka, the rationals) $\mathbb{Q}$ can be constructed as equivalence classes of ordered pairs of integers. Let $(\mathbb {Z} \times (\mathbb {Z} \setminus \{0\}))$⁠ be the set of the pairs $(m, n)$ of integers such that $n \neq 0$ and denoted as $\frac{m}{n}$. An equivalence relation is defined on this set by
\[(m_{1},n_{1})=(m_{2},n_{2})\iff m_{1}n_{2}=m_{2}n_{1}.\]
Addition and multiplication can be defined by the following rules:
\[(m_{1},n_{1})+(m_{2},n_{2})\equiv (m_{1}n_{2}+n_{1}m_{2},n_{1}n_{2}),\]
\[(m_{1},n_{1})\times (m_{2},n_{2})\equiv (m_{1}m_{2},n_{1}n_{2}).\]
\sssc{Real numbers (實數)}
Let $\mathbb{R}$ denote the set of all real numbers. Then:
\bit
\item \tb{Addition and multiplication}: The set $\mathbb{R}$ is a field under addition and multiplication. In other words,
\bit
\item Associativity of addition and multiplication:
\[\forall x,y,z\in\mathbb{R}\colon x+(y+z)=(x+y)+z\land x\cdot(y\cdot z)=(x\cdot y)\cdot z.\]
\item Commutativity of addition and multiplication:
\[\forall x,y\in\mathbb{R}\colon x+y=y+x\land x\cdot y=y\cdot x.\]
\item Distributivity of multiplication over addition:
\[\forall x,y,z\in\mathbb{R}\colon x\cdot(y+z)=(x\cdot y)+(x\cdot z).\]
\item Existence of additive identity:
\[\forall x\in\mathbb{R}\colon x+0=x.\]
\item Existence of multiplicative identity:
\[0\neq 1,\]
\[\forall x\in\mathbb{R}\colon x\times 1=x.\]
\item Existence of additive inverses:
\[\forall x\in\mathbb{R}\colon\exists(-x)\in\mathbb{R}\text{\ s.t.\ }x+(-x)=0.\]
\item Existence of multiplicative inverses:
\[\forall x\in\mathbb{R}\land x\neq 0\colon\exists x^{-1}\in\mathbb{R}\text{\ s.t.\ }x\cdot x^{-1}=1.\]
\eit
\item \tb{Total order}: The field $\mathbb{R}$ is ordered, meaning that there is a total order $\leq$ such that for all real numbers $x$, $y$ and $z$:
\bit
\item Preservation of order under addition:
\[\forall x,y,z\in\mathbb{R}\colon x\leq y\implies x+z\leq y+z.\]
\item Preservation of order under multiplication:
\[\forall x,y\in\mathbb{R}\colon 0\leq x\land 0\leq y\implies 0\leq xy.\]
\item The order is Dedekind-complete, meaning that every nonempty subset $S$ of $\mathbb{R}$ with an upper bound in $\mathbb{R}$ has a least upper bound (aka, supremum) in $\mathbb{R}$. This property applies to the real numbers but not to the rational numbers. For example, $\{x\in\mathbb{Q}\colon x^2<2\}$ has a rational upper bound (e.g., 1.42), but no least rational upper bound, because $\sqrt{2}$ is not rational.
\eit\eit
\sssc{Irrational numbers (無理數)}
\[\mathbb{R}\setminus\mathbb{Q}.\]
\sssc{Complex numbers (複數)}
\[\mathbb{C}=\{a+bi\mid a,b\in\mathbb{R}\land i=\sqrt{-1}\}.\]
\ssc{Almost all (幾乎所有)}
In mathematics, the term "almost all" means "all but a negligible quantity". More precisely, if $X$ is a set, "almost all elements of $X$" means "all elements of $X$ but those in a negligible subset of $X$". The meaning of "negligible" depends on the mathematical context; for instance, it can mean finite, countable, or null.
\ssc{Cover or covering (覆蓋)}
A cover or covering $C$ of a set $X$ is a family of subsets of $X$ such that:
\[\bigcup_{U\in C}U=X.\]
\ssc{Order}
\sssc{Ordered set (序集)}
An ordered set is an ordered pair $P=(X,\leq )$ (or $P=(X,<)$) consisting of a set $X$, called the ground set of $P$, and an order (序) $\leq$ (or $<$), in which an order is a binary relation between two elements in $X$.

A non-strict order $\leq$ is an order such that $a=b\implies a\leq b$; a strict order $<$ is an order such that $a<b\implies a!=b$. An ordered set defined with a non-strict order is called a non-strictly ordered set; an ordered set defined with a strict order is called a strictly ordered set.

Unless otherwise specified,
\bit
\item an order denoted as $\leq$ is a non-strict order;
\item an order denoted as $<$ is a strict order;
\item an order without notation or explanation that makes it a strict order is considered a non-strict order;
\item for an non-strictly ordered set with order $\leq$, $\geq$ is defined as a binary relation between two elements $a,b$ in it such that $a\leq b\iff b\geq a$;
\item for an strictly ordered set with order $<$, $>$ is defined as a binary relation between two elements $a,b$ in it such that $a<b\iff b>a$;
\item for an strictly ordered set with order $<$, a non-strict order $\leq$ is defined as a binary relation between two elements $a,b$ in it such that $a\leq b\iff a<b\lor a=b$; and
\item for a specefic type $P$ of ordered sets (e.g. preordered set), strictly $P$ (e.g. strictly preordered set) is a type of strictly ordered set defined with the same definition of $P$ but replace non-strict order $\leq$ with strict order $<$ and remove the requirement of antisymmetry (if any).
\eit

When the meaning is clear from context and there is no ambiguity about the order, the groud set itself can sometimes represent the whole ordered set.
\sssc{Preordered set (預序集), preset, or quasiordered set}
A preordered set is an ordered pair $P=(X,\leq )$ consisting of a set $X$ and a preorder (預序), aka, quasiorder, $\leq$ on $X$, that is, for all $a,b,c\in X$ it must satisfy:
\bit
\item Reflexivity: $a\leq a$, and
\item Transitivity: $a\leq b\implies a+c\leq b+c$.
\eit
\sssc{Partially ordered set (偏序集) or poset}
A partially ordered set is an ordered pair $P=(X,\leq )$ consisting of a set $X$ and a partial order (偏序) $\leq$ on $X$, that is, for all $a,b,c\in X$ it must satisfy:
\begin{enumerate}
\item Reflexivity: $a\leq a$, i.e. every element is related to itself.
\item Antisymmetry: $a\leq b\land b\leq a\implies a=b$, i.e. no two distinct elements precede each other.
\item Transitivity: $a\leq b\land b\leq c\implies a\leq c$.
\end{enumerate}
\sssc{Upward closure}
Let $A$ be a subset of a poset $X$, the upward closure of $A$, denoted as $\uparrow A$, is defined as:
\[\uparrow A \coloneq \{ x \in X \mid\exists a \in A \text{\ s.t.\ } a \leq x \}.\]
\sssc{Totally ordered set (全序集) or linearly ordered set (線性順序集)}
A totally ordered set is an ordered pair $P=(X,\leq )$ consisting of a set $X$ and a total order (全序), aka, linear order (線性順序), $\leq$ on $X$, that is, for all $a,b,c\in X$ it must satisfy:
\begin{enumerate}
\item Reflexivity: $a\leq a$, i.e. every element is related to itself.
\item Antisymmetry: $a\leq b\land b\leq a\implies a=b$, i.e. no two distinct elements precede each other.
\item Transitivity: $a\leq b\land b\leq c\implies a\leq c$.
\item Strongly connectivity or totality: $a\leq b\lor b\leq a$.
\end{enumerate}
\sssc{Ordered field (有序域)}
A field $(F,+,\cdot)$ together with a total order $\leq$ on $F$ is an ordered field if the order satisfies the following properties for all $a,b,c\in F$:
\begin{enumerate}
\item $a\leq b\implies a+c\leq b+c$.
\item $0\leq a\land 0\leq b\implies 0\leq a\cdot b$.
\end{enumerate}
As usual, we write $a<b$ for $a\leq b$ and $a\neq b$. The notations $b\geq a$ and $b>a$ stand for $a\leq b$ and $a<b$, respectively. Elements $a\in F$ with $a>0$ are called positive.
\sssc{Well ordered set, well-ordered set, or woset (良序集)}
A well ordered set is a totally ordered set such that every non-empty subset of it has a least element in its order. The order of a woset is called a well order or well-order.
\sssc{Order homomorphism or Order-preserving function}
Given two posets $(S,\leq _{S})$ and $(T,\leq _{T})$, an order homomorphism or order-preserving function from $(S,\leq _{S})$ to $(T,\leq _{T})$ is a function $f$ from $S$ to $T$ with the property that, for every $x,y\in S$, $x \leq_S y$ implies $f(x)\leq _{T}f(y)$.
\sssc{Order isomorphism}
Given two posets $(S,\leq _{S})$ and $(T,\leq _{T})$, an order isomorphism from $(S,\leq _{S})$ to $(T,\leq _{T})$ is a bijective function $f$ from $S$ to $T$ with the property that, for every $x,y\in S$, $x \leq_S y$ if and only if $f(x)\leq _{T}f(y)$.

Two sets are called to be order-isomorphic if there exists an order isomorphism between them.
\sssc{Ordered by inclusion}
A family of sets is called to be ordered by inclusion if it is a poset such that for any two sets $A,B$ in it, $A\leq B\iff A\subseteq B$.
\sssc{Cofinal (共尾的) or frequent}
A subset $B\subseteq A$ of a preordered set $(A,\leq )$ is said to be cofinal or frequent in $A$ if for every $a\in A$, there exists a $b\in B$ such that $a\leq b$.
\sssc{Upper Bound (上界)}
$M$ is an upper bound of a preset $A$ if $a\leq M$ for every $a\in A$. We say a preset is bounded-above by $M$ if it has an upper bound $M$.

$M$ is an upper bound of a function $f$ of which the range is a preset if $M$ is an upper bound of its range. We say a function is bounded-above by $M$ if it has an upper bound $M$.

$M$ is an upper bound of a function $f$ of which the range is a preset on a subset $I$ of its domain if $M$ is an upper bound of $f(I)$. We say a function is bounded-above by $M$ on a subset $I$ of its domain if it has an upper bound $M$ on $I$.
\sssc{Lower Bound (下界)}
$m$ is an lower bound of a preset $A$ if $a\geq m$ for every $a\in A$. We say a preset is bounded-below by $m$ if it has an lower bound $m$.

$m$ is an lower bound of a function $f$ of which the range is a preset if $m$ is an lower bound of its range. We say a function is bounded-below by $m$ if it has an lower bound $m$.

$m$ is an lower bound of a function $f$ of which the range is a preset on a subset $I$ of its domain if $m$ is an lower bound of $f(I)$. We say a function is bounded-below by $m$ on a subset $I$ of its domain if it has an lower bound $m$ on $I$.
\sssc{Supremum or Least Upper Bound (最小上界)}
The supremum or least upper bound of a preset $A$, denoted as $\sup A$, is defined as $M_0$ in the set $\mathcal{M}$ of all upper bounds of $A$ such that $\forall M\in\mathcal{M}\colon M_0\leq M$.

The supremum or least upper bound of a function $f$ of which the range is a preset, denoted as $\sup_xf(x)$, is defined as $M_0$ in the set $\mathcal{M}$ of all upper bounds of $f$ such that $\forall M\in\mathcal{M}\colon M_0\leq M$.
\sssc{Infimum or Greatest Lower Bound (最大下界)}
The infimum or greatest lower bound of a preset $A$, denoted as $\inf A$, is defined as $m_0$ in the set $\mathcal{M}$ of all lower bounds of $A$ such that $\forall m\in\mathcal{M}\colon m_0\geq m$.

The infimum or greatest lower bound of a function $f$ of which the range is a preset, denoted as $\inf_xf(x)$, is defined as $m_0$ in the set $\mathcal{M}$ of all lower bounds of $f$ such that $\forall m\in\mathcal{M}\colon m_0\geq m$.
\sssc{Bounded (有界的)}
We say a preset or a function of which the range is a preset is bounded if it is bounded-above and bounded-below.

We say a function is bounded on a subset $I$ of its domain if it is bounded-above and bounded-below on $I$.
\sssc{Directed set (有向集), directed preset (有向預序集), or filtered set}
A directed set is a preset in which every finite subset has an upper bound.
\ssc{Intervals (區間)}
Unless otherwise specified, an interval defined for $\mathbb{R}$.

Right open intervals and left open intervals are collectively called half open intervals (半開區間). All intervals with at least one of the endpoint being $\infty$ or $-\infty$ are collectively called infinite intervals (無限區間).
\sssc{Closed intervals (閉區間) in partially ordered set}
For a partially ordered set $(P,\leq)$ and $a,b\in P$, 
\bit
\item a closed interval $[a,b]$ is defined as
\[[a,b]\coloneq\{x\in P\mid a\leq x\leq b\};\]
\item an left-closed upward interval $[a,\infty)$ is defined as
\[[a,\infty)\coloneq\{x\in P\mid a\leq x\};\]
\item a right-closed downward interval $(-\infty,b]$ is defined as
\[(-\infty,b]\coloneq\{x\in P\mid x\leq b\};\]
\item the interval $(-\infty,\infty)$ is defined as
\[(-\infty,\infty)\coloneq P.\]
\eit
\sssc{Intervals (區間) in strictly partially ordered set}
For a strictly partially ordered set $(P,\leq)$ and $a,b\in P$,
\begin{itemize}
\item an open interval (開區間) $(a,b)$ is defined as
\[(a,b)\coloneq\{x\in P\mid a<x<b\};\]
\item a right open interval (右開區間) $[a,b)$ is defined as
\[[a,b)\coloneq\{x\in P\mid a\leq x<b\};\]
\item a left open interval (左開區間) $(a,b]$ is defined as
\[(a,b]\coloneq\{x\in P\mid a<x\leq b\};\]
\item an left-open upward interval $(a,\infty)$ is defined as
\[(a,\infty)\coloneq\{x\in P\mid a<x\};\]
\item a right-open downward interval $(-\infty,b)$ is defined as
\[(-\infty,b)\coloneq\{x\in P\mid x<b\}.\]
\end{itemize}
\ssc{Ordinal number or ordinal (序數)}
\sssc{Ordinal number or ordinal}
The ordinal number or ordinal is a measurement of the order of an element of a strict well ordered set.

The ordinal numbers is defined to ordered by inclusion to form a strictly well ordered set; the ordinal $0$ is defined as $\varnothing$; and each ordinal $>0$ is defined as the set of all smaller ordinals, that is, for an ordinal $a>0$:
\[a=\{b\mid b\tx{\ is an ordinal and\ }b<a\}.\]
This definition is known as the von Neumann definition of ordinals.

The ordinal number $\omega$, called the first infinite ordinal, is defined to be $\mathbb{N}_0$ in von Neumann definition of ordinals.
\sssc{Successor ordinal}
If an ordinal has a maximum $a$, then it is the next ordinal after $a$, then it is called a successor ordinal (of $a$), written $a+1$.
\sssc{Limit ordinal}
A nonzero ordinal that is not a successor is called a limit ordinal.
\subsection{Cardinality (勢 or 計數) and Cardinal number or cardinal (基數 or 量)}
\sssc{Cardinality and Cardinal number or cardinal}
The cardinal number or cardinal is a measurement of the cardinality (size) of a set $A$, denoted as $|A|$ or $n(A)$.

Two sets are said to be equinumerous or have the same cardinality if there exists a bijection between them.

The cardinality of a finite set is the number of its elements. A cardinal that is not finite is called an infinite cardinal.

Assuming the axiom of choice, the cardinality of a set $X$ is defined to be the least ordinal number $\alpha$ such that there is a bijection between $X$ and $\alpha$. This definition is known as the von Neumann cardinal assignment.
\sssc{Aleph numbers}
Aleph numbers, a system to describe infinite cardinals, denoted as $\aleph_0, \aleph_1, \aleph_2, \ldots \aleph_\omega, \aleph_{\omega+1}, \ldots$, in which the subscripts are ordinal numbers, are defined with:
\bit
\item $\aleph_0$ is defined as $\omega$;
\item for any successor ordinal $n$, $\aleph_n$ is defined inductively as:
\[\aleph_n=\min\{b\mid b\tx{\ is an cardinal and\ }b>\aleph_{n-1}\};\]
\item for any limit ordinal $\lambda$, $\aleph_{\lambda}$ is defined as:
\[\aleph_{\lambda}=\sup\{\aleph_{\alpha}\mid\alpha<\lambda}.\]
\eit
\sssc{Beth numbers}
Beth numbers, a system to describe infinite cardinals, denoted as $\beth_0, \beth_1, \beth_2, \ldots \beth_\omega, \beth_{\omega+1}, \ldots$, in which the subscripts are ordinal numbers, are defined with:
\bit
\item $\beth_0$ is defined as $\aleph_0$;
\item for any successor ordinal $n$, $\beth_n$ is defined inductively as:
\[\beth_n=2^{\aleph_{n-1}};\]
\item for any limit ordinal $\lambda$, $\aleph_{\lambda}$ is defined as:
\[\aleph_{\lambda}=\sup\{\beth_{\alpha}\mid\alpha<\lambda}.\]
\eit
\sssc{Countable infinity (可數無限)}
The cardinality of $\mathbb{N}_0$, called the cardinality of countable infinity (可數無限勢), is $\aleph_0$. A set with the same cardinality of $\mathbb{N}$ is called to be countable infinite (可數無限的).
\sssc{Uncountable infinity (不可數無限)}
The cardinality of $\mathbb{R}$, called the cardinality of continuity (連續勢), is $2^{\aleph_0}=\beth_1$, also denoted as $\mathfrak{c}$. A set with the same cardinality of $\mathbb{R}$ is called to be uncountable infinite (不可數無限的).
\sssc{Continuum hypothesis (CH)}
The continuum hypothesis (CH) states that, there is no set whose cardinality is strictly between that of the integers and the real numbers, that is, $2^{\aleph_0}=\aleph_1$.

CH is independent from ZFC, that is, proving or disproving the CH within ZFC is impossible.
\sssc{Generalized continuum hypothesis (GCH) or Cantor's aleph hypothesis}
The generalized continuum hypothesis (GCH) or Cantor's aleph hypothesis states that, if a set's cardinality lies between that of an infinite set $S$ and that of the power set $P(S)$ of $S$, then it has the same cardinality as either $S$ or $P(s)$, that is, for any infinite cardinal $\lambda$, there is no cardinal $\kappa$ such that $\lambda <\kappa <2^{\lambda }$, that is, $\aleph _{\alpha +1}=2^{\aleph _{\alpha }}$ for every ordinal $\alpha$.

GCH is independent from ZFC, that is, proving or disproving the GCH within ZFC is impossible.
\ssc{Group theory (群論)}
\sssc{Monoid (么半群, 單群, or 亞群)}
A monoid is a set equipped with an operation that combines any two elements of the set to produce a third element within the same set such that the following conditions hold: the operation is associative, and it has an identity element.
\sssc{Group (群)}
A group is a set equipped with an operation that combines any two elements of the set to produce a third element within the same set such that the following conditions hold: the operation is associative, it has an identity element, and every element of the set has an inverse element.
\sssc{Projection maps}
The $k$-th projection map for a Cartesian product of $n$ sets:
\[\prod_{i=1}^nX_i\]
is the function $\pi_k$
\[\pi_k\colon\prod_{i=1}^nX_i\to X_k;\;\pi_k(x_1,x_2,\ldots x_n)=x_k.\]
\end{document}
