\documentclass[a4paper,12pt]{report}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{3}
\input{/usr/share/LaTeX-ToolKit/template.tex}
\begin{document}
\title{Probability Theory}
\author{沈威宇}
\date{\temtoday}
\titletocdoc
\sct{Probability Theory (機率論)}
\subsection{Basic definitions}
\sssc{(Random) Experiment or Trial (（隨機）試驗)}
A (random) experiment or trial is a test that has repeatability, that is, the test can be repeated under the same conditions, and randomness, that is, the results of each test may be different.
\sssc{(Elementary) outcome and sample space (樣本空間)}
The sample space of an experiment is the finest-grain, mutually exclusive, collectively exhaustive set of all possible (elementary) outcomes. For example, the sample space for flipping a coin is $\{\text{head}, \text{tail}\}$.

The singleton of an outcome is sometimes denoted the same as the outcome.
\sssc{Event (事件)}
A subset of the sample space. For example, the event of rolling a die and getting an even number is $\{2, 4, 6\}$.
\sssc{Event space}
The set of all events of an experiment.
\sssc{Probability (機率)}
The likelihood of an event occuring, which is a real number in $[0,1]$. The closer the probability is to 1, the more likely the event is to occur. The probability of an event $A$ occuring is denoted as $P(A)$, $\bbP(A)$, $P[A]$, or $\bbP[A]$.
\sssc{Impossible event (空事件)}
An event with zero probability.
\sssc{Sure event (全事件)}
An event with one probability. If $A$ is a sure event, the trial is said to be almost surely (a.s.) $A$.
\sssc{Sum event (和事件)}
The sum event of event $A$ and event $B$ is $A\cup B$, also denoted as $A+B$.
\sssc{Product event (積事件)}
The product event of event $A$ and event $B$ is $A\cap B$, also denoted as $AB$ or $A,B$.
\sssc{Complement event (餘事件)}
The complement event of event $A$ in sample space $S$ is $S\setminus A$, denoted as $A'$, $A^c$, or $A^C$.
\sssc{Probability space (機率空間)}
A probability space $(\Omega,\calF,P)$ is a measure space of which $\Omega$ is the sample space of an experiment, $\calF$ is the event space of the experiment, and $P$ is the probability measure with unit measure axiom $P(\Omega)=1$ that maps each event to the probability of it.
\sssc{Probability axioms or Kolmogorov axioms}
\begin{enumerate}
\item Nonnegativity:
\[\forall E\in\calF\colon P(E)\geq 0.\]
\item Unit measure or normalization:
\[P(\Omega)=1.\]
\item Countable additivity or \text{\textsigma}-additivity: For all finite or countable infinite indexed family $(E_k)_{k\in K}$ of pairwise disjoint sets in $\calF$,
\[P\left(\bigcup_{k\in K}E_k\right)=\sum_{k=1}^\infty P(E_k).\]
This implies $P(\varnothing)=0$.
\end{enumerate}
\sssc{Partition (分割)}
A partition is a partition of sample space that is a subset of the event space.
\sssc{Classical probability (古典機率) or equally likely outcomes}
If the number of all possible outcomes in the sample space $S$ of an experiment is finite, and the chance of each outcome occurring is equal, then the probability of an event $A$ occurring can be calculated by:
\[P(A)=\frac{|A|}{|S|}.\]
\sssc{Mutually exclusive events (互斥事件)}
Two or more events $A_1,A_2,\ldots,A_n$ are mutually exclusive events iff:
\[\forall i,j\in\bbN_{\le n}\land i\ne j\colon P(A_i\cap A_j)=0.\]
\sssc{(Mutually) independent events (獨立事件)}
Two or more events $A_1,A_2,\ldots,A_n$ are (mutually) independent events iff:
\[\forall J\subseteq\bbN_{\leq n}\land J\neq\varnothing\colon P\qty(\bigcap_{j\in J} A_j)=\prod_{j\in J}P\qty(A_j).\]
\sssc{Sequential experiment}
A sequential experiment is an experiment that
\bit
\item Consists of two or more experiments called subexperiments or stages.
\item Has outcomes at each stage, which may depend on the outcome of the previous stages.
\item The overall outcome of the experiment is an ordered tuple of stage outcomes
\[\omega=(\omega_1,\omega_2,\ldots,\omega_n).\]
\eit
\sssc{Tree diagram}
Tree diagram is often used to represent the possible outcomes, in which the root is the beginning of the experiment, each branch represents a possible stage outcome, and a path from the root to a leaf represents a complete outcome of the sequential experiment. Conditional probabilities can be assigned to branches to compute the probability of each path as the product of the conditional probabilities of all branches in the path.
\sssc{Independent trials}
Independent trials are identical subexperiments in a sequential experiment. The probability space of all the subexperiments are identical and independent of the outcomes of previous subexperiments. Let an event of the sequential experiment be
\[\omega=(\omega_1,\omega_2,\ldots,\omega_n),\]
then
\[P(\omega)=\prod_{i=1}^nP(\omega_i).\]
\sssc{Reliability theory}
A system is in series if all components work iff the system work.

A system is in parallel if at least one component works iff the system work.
\ssc{Counting principle (計數原理) or fundamental principle of counting (基本計數原理)}
\sssc{Complement principle}
\[P(A')=1-P(A).\]
\sssc{De Morgan's laws}
\[P((A\cup B)')=P(A'\cap B').\]
\[P((A\cap B)')=P(A'\cup B').\]
\sssc{Principle of inclusion-exclusion (PIE) or inclusion–exclusion principle (排容原理/取捨原理)}
For event space $\calF$ and indexed family $(A_i)_{i=1}^n\subseteq\calF$ with $n\in\bbN$,
\[P\qty(\bigcup_{i=1}^nA_i)=\sum_{J\subseteq\bbN_{\leq n}\land J\neq\varnothing}(-1)^{|J|+1}P\qty(\bigcap_{j\in J}A_j).\]
\sssc{Subset law}
\[A\subseteq B\implies P(A)\le P(B).\]
\sssc{Union bound}
\[P(\bigcup_{i=1}^nA_i)\le\sum_{i=1}^nP(A_i).\]
\ssc{Independence theorems}
\sssc{Product rule}
If events $A$ and $B$ are independent, $P(A)\cdot P(B)=P(A\cap B)$.
\sssc{Complement independence}
If $A_1,A_2,\ldots,A_n$ are independent, then they, with arbitrary ones replaced with their complement, are still independent.
\sssc{Pairwise and mutal}
Pairwise independence is implied by but do not imply mutual independence.
\sssc{Multinominal distribution}
Suppose a subexperiment has sample space $S=\{s_0,s_1,\ldots,s_{m-1}\}$ with $P(s_i)=p_i$. For $n=\sum_{i=0}^{m-1}n_i$ independent trials, the probability of $n_i$ occurrences of $s_i$ is
\[\binom{n}{n_0,n_1,\ldots,n_{m-1}}\prod_{i=0}^{m-1}p_i^{\pht{i}n_i}.\]
\sssc{Components in series}
Suppose a system $W$ consists of components $W_1,W_2,\ldots,W_n$ in series with probability to work $p_1,p_2,\ldots,p_n$, then the probability for the system to work is
\[P(W)=\prod_{i=1}^np_i.\]
\sssc{Components in parallel}
Suppose a system $W$ consists of components $W_1,W_2,\ldots,W_n$ in parallel with probability to work $p_1,p_2,\ldots,p_n$, then the probability for the system to work is
\[P(W)=1-\prod_{i=1}^n(1-p_i).\]
\ssc{Random variable (RV) (隨機變數)}
\sssc{Random variable}
For a probability spaces $(\Omega_s,\calF_s,P_s)$ and a measurable space $(\Omega_t,\calF_t)$, a random variable $X$ is a measurable function $X\colon\Omega_s\to\Omega_t$ such that for each $E\subseteq\calF_t$, the preimage $X^{-1}(E)\in\calF_s$, and the induced probability spaces $(\Omega_t,\calF_t,P_t)$ of target sample space $\Omega_t$, target event space, and pushforward probability measure $P_t(E)\coloneq P_s(X^{-1}(E))$. It is often the case that $\calF_t$ is the Borel $\text{\textsigma}$-algebra of $\Omega_t$.
\sssc{Real(-valued) random variable}
A random variable is a real(-valued) random variable if $\Omega_t\subseteq\bbR$ and $\calF_t$ is the Borel $\text{\textsigma}$-algebra of $\Omega_t$.
\sssc{Cumulative distribution function (CDF)}
The cumulative distribution function (CDF) of a real random valuable $X$ is defined as
\[F_X(x)\coloneq P(X\le x)=\int_{\Omega_t}\mb{1}_{(-\infty,x]}(t)\dd{P_t(t)}.\]
\sssc{Complementary cumulative distribution function (CCDF)}
The complementary cumulative distribution function (CCDF) of a real random valuable $X$ is defined as
\[\bar{F}_X(x)=1-F_X(x).\]
\sssc{CDF difference theorem}
For $a\ge b$,
\[F_X(b)-F_X(a)=P(X\in(a,b]).\]
\sssc{Non-decreasing property of CDF}
A CDF is non-decreasing.
\bpr
For $x<y$, $\{X|X\le x\}\subseteq\{X|X\le y$, thus $P(X\le x)\le P(X\le y)$.
\epr
\sssc{Limit at negative infinity of CDF}
For a CDF $F_X(x)$, $\lim_{x\to-\infty}F_x(x)=0$.
\bpr
\[x_1>x_2\implies\{X\le x_1\}\supseteq\{X\le x_2\}\]
\[\bigcap_{x\in\bbR}\{X\le x\}=\varnothing\]
By continuity of measure from above,
\[\lim_{x\to-\infty}P(X\le x)=P(\varnothing)=0\]
\epr
\sssc{Limit at positive infinity of CDF}
For a sample space $\Omega$ and CDF $F_X(x)$, $\lim_{x\to\infty}F_x(x)=1$.
\bpr
\[x_1<x_2\implies\{X\le x_1\}\subseteq\{X\le x_2\}\]
\[\bigcup_{x\in\bbR}\{X\le x\}=\Omega\]
By continuity of measure from below,
\[\lim_{x\to\infty}P(X\le x)=P(\Omega)=1\]
\epr
\sssc{Right-continuous property of CDF}
A CDF is right-continuous.
\bpr
\[h_1>h_2\implies\{X\le x+h_1\}\supseteq\{X\le x+h_2\}\]
\[\bigcap_{h>0}\{X\le x+h\}=\{X\le x\}\]
By continuity of measure from above,
\[\lim_{h\to 0^+}P(X\le x+h)=P\qty(\bigcap_{h>0}\{X\le x+h\})=P(X\le x)\]
\epr
\sssc{Converse property of CDF}
If a function $F\colon\bbR\to[0,1]$ is non-decreasing, has limits $0$ at $-\infty$ and $1$ at $\infty$, and is right-continuous, then there exists a probability space $(\Omega,\calF,P)$ and and real random variable $X$ such that $F_X=F$.
\bpr
Let $\Omega=(0,1)$, $\calF=\calB(\Omega)$, and $P$ be Lebesgue measure.

Define
\[X(u)\coloneq\inf\{x\in\bbR\mid F(x)\ge u\},\quad u\in(0,1)\]
\[F_X(x)=F(x)\]
\epr
\sssc{Quantile}
Quantile $Q(p)$ is the inverse of CDF, that is,
\[P(X\le Q(p))=p.\]
\sssc{Discrete random variable}
A discrete random variable is a real random variable with a countable sample space.
\sssc{Continuous random variable}
A continuous random variable is a real random variable with continuous CDF.
\sssc{Simple random variable}
A simple random variable is a discrete random variable $X$ such that there exist a finite partition $\{A_1,\ldots,A_n\}$ of source sample space $\Omega$ such that $X=\sum_{i=1}^na_i\mb{1}_{A_i}$ where $a_i$ are real numbers and $\mb{1}_{A_i}$ is the indicator function of $A_i$, that is,
\[\mb{1}_{A_i}(\omega)=\bcs1,\quad&\omega\in A_i\\0,\quad&\omega\notin A_i\ecs.\]
\sssc{Derived ramdom variable}
For random variables $X_1,\ldots,X_n$ and an $n$-ary function $f$, a random variable $Y=f(X_1,\ldots,X_n)$ is a derived ramdom variable.
\sssc{Independent and identically distributed (i.i.d.) random variables}
Identical random variables that are mutually independent.
\ssc{Discrete Random Variables}
\sssc{Probability mass function (PMF)}
For a discrete random variable $X$ with target sample space $\Omega_t=\{a_1,a_2,\ldots\}$, the probability mass function of $X$, denoted as $p_X$ or $P_X$, is defined as
\[p_X(a_i)=P(X=a_i)\]
for each $a_i\in\Omega_t$.

It has the following properties:
\[p_X(a_i)\ge 0,\quad\forall a_i\in\Omega_t.\]
\[\sum_{a_i\in\Omega_t}p_X(a_i)=1.\]
That the PMF of $X$ is $p$ is denoted as $X\sim p$.
\sssc{PMF determines CDF}
For a discrete random variable $X$ with PMF $p_X$, the CDF $F_X$ can be expressed in terms of the PMF as
\[F_X(x)=\sum_{a_i\le x}p_X(a_i)\]
for all $x\in\bbR$.
\sssc{Support}
The support of a discrete random variable is the support of the PMF.
\sssc{Probability simplex}
Given a sample space $(x_0,x_1,x_2,\ldots,x_k)$, each point $(p_0,p_1,\ldots,p_k)$ in the probability $k$-simplex
\[\{(p_0,p_1,\ldots,p_k)\mid\sum_{i=0}^kp_i=1\}\]
corresponds to the PMF $p$ defined as
\[p(x_i)=p_i,\quad i\in[0,k]\cap\bbZ.\]
\ssc{Continuous Random Variables}
\sssc{Probability density function (PDF)}
For a continuous random variable $X$, the probability density function of $X$, denoted as $f_X$, is defined as
\[f_X(x)=\dv{}{x}F_X(x)\]
for all $x$ such that RHS exists.

It has the following properties:
\[f_X(x)\ge 0.\]
\[\int_{-\infty}^{\infty}f_X(x)\dd{x}=1.\]
That the PDF of $X$ is $p$ is denoted as $X\sim p$.
\sssc{PDF determines CDF}
For a continuous random variable $X$ with PDF $f_X$, the CDF $F_X$ can be expressed in terms of the PMF as
\[F_X(x)=\int_{-\infty}^xf_x(t)\dd{t}\]
for all $x\in\bbR$.
\sssc{Support}
The support of a discrete random variable is the support of the PDF.
\ssc{Sum random variable}
\sssc{General case}
For two independent real random variables $X,Y$ with the same source sample space and event space, and with pushforward probability measure $P_X$ and $P_Y$ respectively, let the Schwartz distributions induced by $P_X$ and $P_Y$ be $T_X$ and $T_Y$ respectively. Then the Schwartz distribution $T_{X+Y}$ induced by the pushforward probability measure $P_{X+Y}$ of the sum random variable $X+Y$ is the distributional convolution of $T_1$ and $T_2$, that is,
\[T=T_1*T_2.\]
\sssc{For discrete random variables}
For two independent discrete random variables $X,Y$,
\[p_{X+Y}(z)=\sum_{x\in\bbR}p_X(x)p_Y(z-x).\]
\sssc{Sum of independent continuous random variables}
For two independent continuous random variables $X,Y$,
\[f_{X+Y}(z)=\int_{-\infty}^{\infty}f_X(x)f_Y(z-x)\dd{x}.\]
\ssc{Conditional Probability (條件機率)}
\sssc{Conditional probability}
The probability of another event $A$ occurring given that a certain event $B$ has occurred, denoted as $P(A|B)$.
\[P(A|B)= \frac{P\qty(A\cap B)}{P\qty(B)}.\]
\sssc{Conditional probability space}
For a probability space $(\Omega,\calF,P)$ and a fixed event $B\in\calF$ with $P(B)>0$, the function $P(\cdot|B)\colon\calF\to[0,1]$ defined by $A\mapsto P(A|B)$ is a probability measure on the same measurable space $(\Omega,\calF)$.
\sssc{Conditional random variable}
For a probability space $(\Omega_s,\calF_s,P_s)$, a measurable space $(\Omega_t,\calF_t)$, a random variable $X\colon\Omega_s\to\Omega_t$ with pushforward probability measure $P_t$, and a fixed event $A\in\calF_t$ with $P_t(A)>0$, the condition random variable $X|B$ is a randm variable over $(\Omega_s,\calF_s,P_s(\cdot|X^{-1}(A)))$ with pushforward probability measure $P_t(\cdot|B)$.
\sssc{Conditional CDF}
For a real random variable $X$ and an event $A\subseteq X(\Omega)$ such that $P(A)>0$, the conditional CDF of $X$ given $A$ follows
\[F_{X|A}(x)=P(X\le x|A)=\frac{P(X\le x,A)}{P(A)},\quad\forall x\in\bbR.\]
\sssc{Conditional PMF of discrete random variable}
For a discrete random variable $X$ and an event $A\subseteq X(\Omega)$ such that $P(A)>0$, the conditional PMF of $X$ given $A$ follows
\[p_{X|A}(x)=P(X=x|A)=\frac{P(X=x,A)}{P(A)}=\frac{P(X=x)\cdot\mb{1}_A(x)}{P(A)},\quad\forall x\in\bbR.\]
\sssc{Conditional PDF of continuous random variable}
For a continuous random variable $X$ and an event $A\subseteq X(\Omega)$ such that $P(A)>0$, the conditional PDF of $X$ given $A$ follows
\[f_{X|A}(x)=\dv{}{x}F_{x|A}(x)=\dv{}{x}(\frac{P(X\le x,A)}{P(A)})=\frac{f_X(x)\cdot\mb{1}_A(x)}{P(A)},\quad\forall x\in\bbR.\]
\ssc{Conditional probability laws}
\sssc{Law of multiplication}
For sample space $\Omega$ and $n$ events $A_1,A_2,\ldots,A_n$, define $B_i=\bigcap_{j=1}^{i-1}A_j$ for $i\in\bbN_{\ge 2,\le n+1}$ and $B_1=\Omega$. Then,
\[P(\bigcap_{i=1}^nA_i)=\bcs\prod_{i=1}^nP(A_i|B_i),\quad&\forall i\in\bbN_{\le n}\colon P(B_i)>0\\0,\quad&\exists i\in\bbN_{\le n}\colon P(B_i)=0\ecs.\]
\bpr
\[\prod_{i=1}^nP(A_i|B_i)=\prod_{i=1}^n\frac{P(B_{i+1})}{P(B_i)}=\frac{P(B_{n+1})}{P(B_1)}=\frac{P(\bigcap_{i=1}^nA_i)}{1}=P(\bigcap_{i=1}^nA_i).\]
\epr
\sssc{Law of total probability}
Let $(A_i)_{i=1}^n\subseteq\calF$ be a partition, and $B$ be an event. Then
\[P(B)=\sum_{i=1}^nP(A_i\cap B)=\sum_{i=1}^nP(B|A_i)P(A_i).\]
Let $X$ be a discrete random variable. Then
\[p_X(x)=\sum_{i=1}^np_{X|A_i}(x)P(A_i).\]
Let $X$ be a continuous random variable. Then
\[f_X(x)=\sum_{i=1}^nf_{X|A_i}(x)P(A_i).\]
\sssc{Bayes' theorem (貝葉斯定理或貝氏定理)}
Let $(A_i)_{i=1}^n\subseteq\calF$ be a partition, and $B$ be an event. Then,
\[P\qty(A_k\mid B)=\frac{P\qty(B|A_k)P(A_k)}{\sum_{i=1}^nP\qty(B\mid A_i)P(A_i)}.\]
\ssc{Multivariate or joint random variable}
\sssc{Definition}
For $n$ real random variables $X_1,X_2,\ldots,X_n$ with the same source sample space $\Omega$ and event space $\calF$, the multivariate or joint random variable $X_1,X_2,\ldots,X_n$ is defined for all $\omega\in\Omega^n$ as
\[X_1,X_2,\ldots,X_n(\omega)=(X_1(\omega),X_2(\omega),\ldots,X_n(\omega)).\]
\sssc{Cumulative distribution function (CDF)}
For $n$ real random variables $X_1,X_2,\ldots,X_n$ with the same source sample space $\Omega$ and event space $\calF$,
\[F_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)=P(X_1\le x_1,X_2\le x_2,\ldots,X_n\le x_n).\]
$X_1,X_2,\ldots,X_n$ are independent iff
\[F_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)=\prod_{i=1}^nF_{X_i}(x_i).\]
\sssc{Probability mass function (PMF)}
For discrete random variables $X_1,X_2,\ldots,X_n$ with the same source sample space $\Omega$ and event space $\calF$,
\[\ba
p_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)&=P(X_1=x_1,X_2=x_2,\ldots,X_n=x_n)\\
&=P(X_1=x_1)\cdot P(X_2=x_2\mid X_1=x_1)\cdot P(X_3=x_3\mid X_1=x_1,X_2=x_2)\cdot\\
&\ldots\cdot P(X_n=x_n\mid X_1=x_1,X_2=x_2\ldots X_{n-1}=x_{n-1})
\ea\]
If $X_1,X_2,\ldots,X_n$ are independent
\[p_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)=\prod_{i=1}^np_{X_i}(x_i).\]
\sssc{Probability density function (PDF)}
For continuous random variables $X_1,X_2,\ldots,X_n$ with the same source sample space $\Omega$ and event space $\calF$,
\[f_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)=\frac{\partial^nF_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)}{\partial x_1\partial x_2\ldots\partial x_n}.\]
If $X_1,X_2,\ldots,X_n$ are independent
\[f_{X_1,X_2,\ldots,X_n}(x_1,x_2,\ldots,x_n)=\prod_{i=1}^nf_{X_i}(x_i).\]
\ssc{Discrete to continuous limit}
We say a sequence or net $\langle X_n\rangle$ of discrete random variables converge to a continuous random variable $X$ if $F_{X_n}$ converges pointwise to $F_X$.
\ssc{(Mathematical) Expected Value, Expectation (期望值), or Mean (平均值)}
\sssc{Definition}
For a random variable $X$ on $(\Omega_s,\calF_s,P_s)$ with induced probability space $(\Omega_t,\calF_t,P_t)$, the expected value or expectation $E(X)$, $\bbE(X)$, $E[X]$, or $\bbE[X]$, or mean $\mu_X$, of it is defined as the Lebesgue integral
\[E(X)=\int_{\Omega_s}X(\omega_s)\dd{P_s(\omega_s)}\]
or equivalently,
\[E(X)=\int_{\Omega_t}x\dd{P_t(x)},\]
or in Stieltjes integral
\[E(X)=\int_{x=-\infty}^{\infty}x\dd{F_X(x)},\]
\sssc{For discrete random variable}
For a discrete random variable $X$ with target sample space $\Omega_t$ and probability mass function $p_X(x)$, the expected value or expectation $E(X)$, $\bbE(X)$, $E[X]$, or $\bbE[X]$, or mean $\mu_X$, of it is
\[E(X)=\sum_{x\in\Omega_t}xp_X(x)\]
if $\Omega_t$ is a finite set or $\Omega_t$ is a countably infinite set and
\[\sum_{x\in\Omega_t}\abs{x}p_X(x)\]
is convergent, and doesn't exist if $\Omega_t$ is a countably infinite set and
\[\sum_{x\in\Omega_t}\abs{x}p_X(x)\]
is divergent.
\sssc{For continuous random variable}
For a continuous random variable $X$ with probability density function $f_X(x)$, the expected value or expectation $E(X)$, $\bbE(X)$, $E[X]$, or $\bbE[X]$, or mean $\mu_X$, of it is
\[E(X)=\int_{-\infty}^{\infty}xf_X(x)\dd{x}\]
if it converges absolutely.
\sssc{Derived random variable theorem}
\[E(f(X))=\int_{\Omega_t}f(x)\dd{P_t(x)}\]
if $f\in L^1(\Omega_t,P_t)$.
\sssc{Expected value from CDF}
For a non-negative real random variable $X$,
\[E(X)=\int_0^{\infty}P(X\ge x)\dd{x}=\int_0^{\infty}1-F_X(x)\dd{x}.\]
\bpr
\[\ba
E(X)&=\int_{\Omega_t}x\dd{P_t(x)}\\
&=\int_{\Omega_t}\int_0^x1\dd{t}\dd{P_t(x)}\\
&=\int_0^{\infty}\int_{\Omega_t}\mb{1}_{t\in[0,x)}(x,t)\dd{P_t(x)}\dd{t}\\
&=\int_0^{\infty}P(X\ge t)\dd{t}
\ea\]
\epr
For real random variable $X$, let
\[X^+=\bcs X,\quad&X>0\\0,\quad&X\le 0\ecs\]
and
\[X^-=\bcs -X,\quad&X<0\\0,\quad&X\ge 0\ecs.\]
Then,
\[E(X^+)=\int_0^{\infty}P(X\ge x)\dd{x}=\int_0^{\infty}1-F_X(x)\dd{x},\]
\[E(X^-)=\int_{-\infty}^0P(X\le x)\dd{x}=\int_{-\infty}^0F_X(x)\dd{x},\]
\bpr
\[\ba
E(X^-)&=\int_0^{\infty}P(X^-\ge u)\dd{u}\\
&=\int_0^{\infty}P(X\le -u)\dd{u},\quad x=-u\\
&=\int_{-\infty}^0P(X\le x)\dd{x}\\
&=\int_{-\infty}^0F_X(x)\dd{x}
\ea\]
\epr
and
\[E(X)=E(X^+)-E(X^-).\]
\sssc{Affine Transformation}
For a real random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[E(aX+b)=aE(X)+b.\]
\sssc{Linearity}
For indexed families of real random variables $(X_i)_{i\in I}$ and real numbers $(a_i)_{i\in I}$ with $\abs{I}\in\bbN\cup\{\aleph_0\}$, where for $\abs{I}=\aleph_0$ suppose RHS is convergent,
\[E(\sum_{i\in I}a_iX_i)=\sum_{i\in I}a_iE(X_i).\]
\sssc{Conditional expectation}
\[E(X|A)=\frac{E(X\cdot\mb{1}_A)}{P(A)}\]
\sssc{Law of total expectation}
Let $(A_i)_{i=1}^n\subseteq\calF$ be a partition. Then
\[E(X)=\sum_{i=1}^nE(X|A_i)P(A_i)\]
\ssc{Variance and Mean square error (MSE)}
\sssc{Variance (變異數/方差)}
For a real random variable $X$, variance $\Var(X)$ or $\sigma_X^{\phantom{X}2}$ of it is defined as
\[\Var(X)=E((X-E(X))^2)=E(X^2)-2E(XE(X))+E(X)^2=E(X^2)-2E(X)^2+E(X)^2=E(X^2)-E(X)^2.\]
\sssc{Standard Deviation (標準差)}
For a real random variable $X$, standard deviation $\SD(X)$ or $\sigma_X$ of it is defined as
\[\SD(X)=\sqrt{\Var(X)}.\]
\sssc{Zero variance}
Variance of a real random variable $X$ is zero iff $X$ is almost surely a constant, that is, there exists a constant $c$ such that $P(X=c)=1$.
\sssc{Covariance (共變異數/共方差)}
For two real random variables $X,Y$, their covariance $\Sigma$ or $\Cov(X,Y)$ is defined as
\[\Cov(X,Y)=E\qty((X-E(X))(Y-E(Y)))=E(X)E(Y)-E(XY).\]
For an $n$-dimensional vector of real random variables $\mb{X}=(X_1,X_2,\ldots)$ with $n\in\bbN\cup\{\aleph_0\}$, the covariance matrix $\Sigma$ or $\Cov(\mb{X})$ is defined as
\[\Sigma=\Cov(\mb{X})=E\qty((\mb{X}-E(\mb{X}))(\mb{X}-E(\mb{X}))^{\top}),\]
that is,
\[\Sigma_{ij}=\Cov(X_i,X_j)=E\qty((X_i-E(X_i))(X_j-E(X_j))).\]
\sssc{Independent random variables covariance matrix}
If all $X_i$ are independent, then $\Cov(\mb{X})$ is a diagonal matrix where $\Sigma_{ii}=\Var(X_i)$.
\sssc{Sum of independent random variables}
For $n$-dimensional vectors of real random variables $\mb{X}=(X_1,X_2,\ldots,X_n)$ and real numbers $\mb{a}=(a_1,a_2,\ldots,a_n)$ with $n\in\bbN$,
\[\Var(\sum_{i=1}^na_iX_i)=\mb{a}\Cov(\mb{X})\mb{a}^{\top}.\]
For real random variables $X,Y$ and real numbers $a,b$,
\[\Var(X+Y)=\Var(X)+\Var(Y)+2\Cov(X,Y).\]
\sssc{Affine Transformation}
For a real random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[\Var(aX+b)=a^2\Var(X).\]
\[\SD(aX+b)=|a|\SD(X).\]
\sssc{Mean square error (MSE) (均方（誤）差) or Expected square error}
For a real random variable $X$ and a prediction or (blind) estimate $\hat{x}$, the mean square error $\MSE_X(\hat{x})$ or expected square error of the prediction is defined as
\[\MSE_X(\hat{x})=E\qty((X-\hat{x})^2).\]
\sssc{Minimum mean square error}
The minimum mean square error estimate of a real random variable $X$ is $E(X)$, and the MSE of it is $\Var(X)$.
\ssc{Moment}
\sssc{Moment}
The $n$th moment of $X$ is defined as $E(X^n)$ for $n\in\bbN$. The zero-th moment $E(X^0)$ is defined as $1$.
\sssc{Central moment}
The $n$th central moment of $X$ is defined as $E((X-E(X))^n)$. The zero-th central moment $E((X-E(X))^0)$ is defined as $1$.
\sssc{Moments determine finite-range random variable theorem}
A random variable $X$ with finite range is uniquely determined by the zero-th to $(n-1)$th moments $\mu_0=1,\mu_1,\ldots,\mu_{n-1}$.
\bpr
Let the range of $X$ be $\{x_1,x_2,\ldots,x_n\}$, moments be $\{\mu_0=1,\mu_1,\ldots,\mu_{n-1}\}$, and $P(X=x_i)=p_i$. Then
\[\begin{bmatrix}
1 &  1 & \ldots & 1 \\
x_1 & x_2 & \ldots & x_n \\
x_1^{\pht{1}2} & x_2^{\pht{2}2} & \ldots & x_n^{\pht{n}2} \\
\vdots & \vdots & \ddots & \vdots \\
x_1^{\pht{1}n-1} & x_2^{\pht{2}n-1} & \ldots & x_n^{\pht{n}n-1}
\end{bmatrix}
\begin{bmatrix}
p_1 \\ p_2 \\ \vdots \\ p_n
\end{bmatrix}
=
\begin{bmatrix}
\mu_0 \\ \mu_1 \\ \vdots \\ \mu_{n-1}
\end{bmatrix}\]
\epr
\sssc{Moment generating function (MGF)}
The moment generating function (MGF) of a random variable $X$ is defined as
\[M_X(t)=E\qty(e^{tX}),\quad t\in\bbR.\]
\sssc{MGF as Laplace transform}
In Stieltjes integral,
\[M_X(t)=\int_{x=-\infty}^{\infty}e^{tx}\dd{F_x(x)}.\]
\sssc{MGF determines CDF}
If the MGF of $X$ exists in a neighborhood of $t=0$, it uniquely determines the distribution of $X$.
\bpr
By Mellin's inverse formula.
\epr
\sssc{MGF generating moments}
If the MGF of $X$ exists in a neighborhood of $t=0$, $E(X^n)=M_X^{\pht{X}(n)}(0)$.
\bpr
Suppose the MGF of $X$ exists in $(-h,\vepsilon)$ for some $\vepsilon>0$ and $J=\left(-\frac{h}{2},\frac{h}{2}\right)$.

For any $\delta>0$ and $n\in\bbN_0$, there exists $C_n>0$ such that
\[|X|^n\le C_ne^{\delta|X|}\]
Take some $\delta$ such that $0<\delta<\frac{h}{2}$,
\[\abs{\dv[n]{}{t}e^{tX}}=|X|^ne^{tX}\le C_ne^{(\delta+t)|X|}\le C_ne^{\frac{h}{2}|X|}\le C_n\qty(e^{\frac{h}{2}X}+e^{-\frac{h}{2}X})\]
Thus, $\dv[n]{}{t}e^{tX}$ is uniformly bounded in $J$.

By Leibniz integral rule, since the integrands are uniformly bounded and thus the integrals are finite, for $t\in J$,
\[\dv{}{t}E\qty(e^{tX})=\dv{}{t}\int_{\bbR}e^{tX}\dd{P(X)}=\int_{\bbR}\dv{}{t}e^{tX}\dd{P(X)}=E(\dv{}{t}e^{tX})\]
Apply Leibniz integral rule iteratively, we obtain
\[\dv[n]{}{t}E\qty(e^{tX})=E\qty(\dv[n]{}{t}e^{tX})\]
Thus,
\[M_X^{\pht{X}(n)}(t)=E(X^ne^{tX})=\sum_{k=n}^{\infty}\frac{t^{k-n}}{(k-n)!}E(X^k)\]
\[M_X^{\pht{X}(n)}(0)=E(X^k)\]
\epr
\sssc{Affine Transformation}
For a real random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[M_{aX+b}(t)=e^{bt}M_X(at).\]
\sssc{Logarithmic Linearity}
For indexed families of real random variables $(X_i)_{i\in I}$ and real numbers $(a_i)_{i\in I}$ with $\abs{I}\in\bbN\cup\{\aleph_0\}$, where for $\abs{I}=\aleph_0$ suppose RHS is convergent,
\[M_{\sum_{i\in I}a_iX_i}(t)=\prod_{i\in I}M_{X_i}\qty(a_it).\]
\ssc{Mode (眾數)}
\sssc{For discrete random variable}
For a discrete random variable $X$ with PMF $p_X(x)$, $\Mode(X)$ is defined as
\[\Mode(X)=\arg\max_xp_X(x),\]
which is not necessarily unique.
\sssc{For continuous random variable}
For a continuous random variable $X$ with PDF $f_X(x)$, $\Mode(X)$ is defined as
\[\Mode(X)=\arg\max_xf_X(x),\]
which is not necessarily unique.
\sssc{Affine Transformation}
For a real random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[\Mode(aX+b)=a\Mode(X)+b.\]
\ssc{Median (中位數)}
\sssc{Definition}
For a real variable \( X \), the median 
$\Median(X)$ of it is define by
\[P(X\leq\Median(X))\geq\frac{1}{2}\land P(X\geq\Median(X))\geq\frac{1}{2},\]
which is not necessarily unique.

When we denote $\Median(X)$ to be an interval $I$, it means that for all $x\in I$,
\[P(X\leq x)\geq\frac{1}{2}\land P(X\geq x)\geq\frac{1}{2}.\]
\sssc{Affine Transformation}
For a real random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[\Median(aX+b)=a\Median(X)+b.\]
\ssc{Entropy (熵) for discrete random variables}
\sssc{Definition}
For a discrete random variable $X$ with target sample space $\Omega$ and PMF $p_X(x)$, the entropy $H(X)$ of base $b$ is defined as
\[H(X)=E(-\log_b(p_X(x)))=-\sum_{x\in\Omega\land p_X(x)>0}p_X(x)\log_b(p_X(x)),\]
where base $b=2$ gives the unit of bits or shannons, base $b=e$ gives natural units or nat, and base $b=10$ gives the unit of dits, bans, or hartleys. When not specified, $b=e$ is used.
\sssc{Nonnegativity}
\[H(X)\ge 0.\]
\bpr
If $p_X(x)=1$ for some $x$, $H(X)=0$; otherwise, for all $x\in\Omega$ such that $p_X(x)>0$, $p_X(x)\in(0,1)$, $\log_b(p_X(x))\in(-\infty,0)$, and $p_X(x)\log(p_X(x))<0$.
\epr
\sssc{Maximum entropy}
If $\abs{\Omega}=n\in\bbN$,
\[H(X)\le\log_bn,\]
and $H(X)=\log_bn$ occurs when $p_X(x)=\frac{1}{n}$ for all $x\in\Omega$.
\bpr
\[H(X)=-\sum_{x\in\Omega\land p_X(x)>0}p_X(x)\log_b(p_X(x))\]
For each $x_i\in\Omega$, let $p_i=p_X(x_i)$,
\[H(X)=-\sum_{i=1}^np_i\log_b(p_i)\]
\[\pdv{H(X)}{p_i}=-\log_b(p_i)-1\]
\[\sum_{i=1}^np_i=1\]
\[\pdv{\sum_{i=1}^np_i-1}{p_i}=1\]
\[-\log_b(p_i)-1-\lambda=0\]
\[p_i=b^{-1-\lambda}\lor0\]
Suppose $m$ out of $n$ $p_i$ are not $0$,
\[H(X)=-mb^{-1-\lambda}(-1-\lambda)\]
Thus when $m=n$, $H(X)$ has maximum, and $p_i=\frac{1}{n}$, and
\[H(X)=-n\frac{1}{n}(-\log_bn)=\log_bn\]
\epr
\sssc{Affine Transformation}
For a discrete random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[H(aX+b)=\begin{cases}H(X),\quad &a\neq 0\\0,\quad &a=0\end{cases}.\]
\ssc{(Differential or continuous) entropy for continuous random variables}
\sssc{Definition}
For a continuous random variable $X$ with target sample space $\Omega$ and PDF $f_X(x)$, the (differential or continuous) entropy $h(X)$ or $H(X)$ of base $b$ is defined as
\[h(X)=E(-\log_b(f_X(x)))=-\int_{\Omega}f(x)\log_b(f(x))\dd{x},\]
where base $b=2$ gives the unit of bits or shannons, base $b=e$ gives natural units or nat, and base $b=10$ gives the unit of dits, bans, or hartleys. When not specified, $b=e$ is used. Differential entropy can be positive.
\sssc{Affine Transformation}
For a continuous random variable $X$, $a\in\bbR_{\neq0}$, and $b\in\mathbb{R}$,
\[h(Y)=h(X)+\log|a|.\]
\subsection{Families of Discrete Random Variables}
\sssc{Discrete uniform random variable}
\bit
\item Notation: $X\sim\calU\{a,b\}$, $X\sim\operatorname{Unif}\{a,b\}$, $X\sim\operatorname{Uniform}\{a,b\}$, $X\sim\calU\{a,a+1,\ldots,b\}$, $X\sim\operatorname{Unif}\{a,a+1,\ldots,b\}$, or $X\sim\operatorname{Uniform}\{a,a+1,\ldots,b\}$, read as $X$ is discretely uniform distributed with parameters $a$ and $b$
\item Parameters: $a,b\in\bbZ\land a\le b$
\item Support: $\bbZ\cap[a,b]$
\item PMF:
\[p_X(k)=\frac{1}{b-a+1},\quad k\in\bbZ\cap[a,b]\]
\item CDF:
\[F_X(k)=\bcs0,\quad&k<a\\\frac{\lfloor k\rfloor-a+1}{b-a+1},\quad&a\le k<b\\1,\quad&k\ge b\ecs\]
\item Mean:
\[E(X)=\frac{a+b}{2}\]
\item Variance:
\[\operatorname{Var}(X)=\frac{(b-a+1)^2-1}{12}\]
\bpr
\[\ba
E(X^2)&=\sum_{k=a}^b\frac{k^2}{b-a+1}\\
&=\frac{1}{b-a+1}\sum_{k=0}^{b-a}(a+k)^2\\
&=\frac{1}{b-a+1}\qty((b-a+1)a^2+2a\frac{(b-a+1)(b-a)}{2}+\frac{(b-a+1)(b-a)(2b-2a+1)}{6})\\
&=a^2+a(b-a)+\frac{(b-a)(2b-2a+1)}{6}\\
&=ab+\frac{2b^2-2ab+b-2ab+2a^2-a}{6}\\
&=\frac{2a^2+2b^2+2ab-a+b}{6}
\ea\]
\[\ba
\operatorname{Var}(X)&=E(X^2)-E(X)^2\\
&=\frac{2a^2+2b^2+2ab-a+b}{6}-\frac{a^2+b^2+2ab}{4}\\
&=\frac{4a^2+4b^2+4ab-2a+2b-3a^2-3b^2-6ab}{12}\\
&=\frac{a^2+b^2-2ab-2a+2b}{12}\\
&=\frac{(b-a+1)^2-1}{12}
\ea\]
\epr
\item MGF:
\[M_X(t)=\frac{e^{at}-e^{(b+1)t}}{(b-a+1)(1-e^t)}\]
\bpr
\[E(e^{tX})=\frac{1}{b-a+1}\sum_{k=a}^be^{kt}=\frac{1}{b-a+1}e^{at}\frac{1-e^{(b-a+1)t}}{1-e^t}=\frac{e^{at}-e^{(b+1)t}}{(b-a+1)(1-e^t)}\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=\bbZ\cap[a,b]\]
\item Median:
\[\operatorname{Median}(X)=\bcs\frac{a+b}{2},\quad&\frac{b-a}{2}\in\bbZ\\\qty[\frac{a+b-1}{2},\frac{a+b+1}{2}],\quad&\frac{b-a+1}{2}\in\bbZ\ecs\]
\item Entropy:
\[H(X)=-\ln(b-a+1)\]
\bpr
\[\ba
H(X)&=-\sum_{k=a}^b\frac{1}{b-a+1}\ln(\frac{1}{b-a+1})\\
&=-\ln(b-a+1)
\ea\]
\epr
\item PMF monotonicity: constant
\eit
\sssc{Bernoulli (伯努力) random variable}
\bit
\item Trial: A Bernoulli random variable models a single Bernoulli trial (aka binary trial) with exactly two possible outcomes, success or 1, with probability $p$, and failure or 0, with probability $1-p$.
\item Notation: $X\sim\operatorname{Bern}(p)$, read as $X$ is Bernoulli distributed with parameter $p$
\item Parameters: $p\in[0,1]$
\item Support: $\{0,1\}$
\item PMF:
\[p_X(k)=\bcs p,\quad&k=1\\1-p,\quad&k=0\ecs\]
\item CDF:
\[F_X(k)=\bcs 0,\quad&k<0\\1-p,\quad&0\le k<1\\1,\quad&k\ge 1\ecs\]
\item Mean:
\[E(X)=p\]
\item Variance:
\[\operatorname{Var}(X)=p(1-p)\]
\bpr
\[E(X^2)=p\]
\[\operatorname{Var}(X)=p-p^2=p(1-p)\]
\epr
\item MGF:
\[M_X(t)=pe^t+1-p\]
\item Mode:
\[\operatorname{Mode}(X)=\bcs0,\quad&p<\frac{1}{2}\\0,1,\quad&p=\frac{1}{2}\\1,\quad&p>\frac{1}{2}\ecs\]
\item Median:
\[\operatorname{Median}(X)=\bcs
0,\quad&p<\frac{1}{2}\\
[0,1],\quad&p=\frac{1}{2}\\
1,\quad&p>\frac{1}{2}
\ecs\]
\item Entropy:
\[H(X)=-p\ln p-(1-p)\ln(1-p)\]
\item PMF monotonicity: strictly increasing if $p>\frac{1}{2}$, constant if $p=\frac{1}{2}$, and strictly decreasing if $p<\frac{1}{2}$
\eit
\sssc{Binomial random variable}
\bit
\item Trial: A binomial random variable counts the number of successes in $n$ Bernoulli trials with success probability $p$.
\item Notation: $X\sim\operatorname{Bin}(n,p)$, $X\sim B(n,p)$, $X\sim\operatorname{Binom}(n,p)$, or $X\sim\operatorname{Binomial}(n,p)$, read as $X$ is binomial distributed with parameters $n$ and $p$
\item Parameters: $n\in\bbN$, $p\in[0,1]$
\item Support: $\bbZ\cap[0,n]$
\item PMF:
\[p_X(k)=\binom{n}{k}p^k(1-p)^{n-k},\quad k\in\bbZ\cap[0,n]\]
\item CDF:
\[F_X(k)=\bcs 0,\quad&k<0\\\sum_{i=0}^{\lfloor k\rfloor}\binom{n}{i}p^i(1-p)^{n-i},\quad&0\le k<n\\1,\quad&k\ge n\ecs\]
\item Mean:
\[E(X)=np\]
\bpr
\[\ba
E(X)&=\sum_{k=1}^nk\binom{n}{k}p^k(1-p)^{n-k}\\
&=\sum_{k=1}^nn\binom{n-1}{k-1}p^k(1-p)^{n-k}\\
&=np\sum_{k=1}^n\binom{n-1}{k-1}p^{k-1}(1-p)^{n-k}\\
&=np(p+1-p)^{n-1}\\
&=np
\ea\]
Alternatively, decompose it to $n$ i.i.d. Bernoulli random variable with parameter $p$. Thus, the mean of $X$ is $n$ times that of the Bernoulli random variable.
\epr
\item Variance:
\[\operatorname{Var}(X)=np(1-p)\]
\bpr
\[\ba
E(X^2)&=\sum_{k=0}^nk^2\binom{n}{k}p^k(1-p)^{n-k}\\
&=np\sum_{k=1}^nk\binom{n-1}{k-1}p^{k-1}(1-p)^{n-k}\\
&=np\sum_{k=1}^{n-1}k\binom{n-1}{k}p^k(1-p)^{n-1-k}+np\\
&=np(n-1)p+np\\
&=n^2p^2-np^2+np
\ea\]
\[\operatorname{Var}(X)=E(X^2)-E(X)^2=-np^2+np=np(1-p)\]
Alternatively, use decomposition.
\epr
\item MGF:
\[M_X(t)=\qty(pe^t+1-p)^n\]
\item Mode:
\[\operatorname{Mode}(X)=\bcs n,\quad&p=1\\(n+1)p-1,(n+1)p,\quad&(n+1)p\in\bbN\land p\notin\bbZ\\\lfloor(n+1)p\rfloor,\quad&\text{otherwise}\ecs\]
\bpr
If $p=0$, $p_X(k)=0$. If $p=1$, $p_X(k)=n$.
Let
\[K\coloneq\arg\max_kp_X(k)\]
Let
\[g(k)\coloneq\frac{p_X(k+1)}{p_X(k)}=\frac{\binom{n}{k+1}p^{k+1}(1-p)^{n-k-1}}{\binom{n}{k}p^k(1-p)^{n-k}}=\frac{(n-k)p}{(k+1)(1-p)}\]
Extend $g(k)$ to the $\bbR\cap[0,n]$. For $k\in(0,n)$,
\[\ba
g'(k)&=\frac{-(k+1)-(n-k)}{(k+1)^2}\frac{p}{1-p}\\
&=\frac{-n+1}{(k+1)^2}\frac{p}{1-p}\\
&\le 0
\ea\]
\[K=1+\min\{k\mid g(k)>1\},1+\min\{k\mid g(k)\ge1\}\]
Solve $g(k*)=1$ for $k*$.
\[\frac{n-k*}{k*+1}=\frac{1-p}{p}\]
\[np-k*p=k*+1-k*p-p\]
\[k*=(n+1)p-1\]
\epr
\item PMF monotonicity: strictly increasing for $k\le(n+1)p$ and strictly decreasing for $k\ge(n+1)p$
\item Decomposition: sum of $n$ i.i.d. Bernoulli random variable with parameter $p$
\eit
\sssc{Geometric random variable}
\bit
\item Trial: A geometric random variable counts the number of independent Bernoulli trials with success probability $p$ until the first success.
\item Notation: $X\sim\operatorname{Geom}(p)$, $X\sim G(p)$, or $X\sim\operatorname{Geometric}(p)$, read as $X$ is geometric distributed with parameter $p$
\item Parameters: $p\in(0,1]$
\item Support: $\bbN$
\item PMF:
\[p_X(k)=p(1-p)^{k-1},\quad k\in\bbN\]
\item CDF:
\[F_X(k)=\bcs 0,\quad&k<1\\1-(1-p)^{\lfloor k\rfloor},\quad&k\ge1\ecs\]
\bpr
\[\ba
P(X\le k)&=\sum_{i=1}^{\lfloor k\rfloor}p(1-p)^{i-1}\\
&=\sum_{i=1}^{\lfloor k\rfloor}p(1-p)^{i-1}\\
&=\sum_{i=1}^{\infty}p(1-p)^{i-1}-\sum_{i=\lfloor k\rfloor+1}^{\infty}p(1-p)^{i-1}\\
&=1-(1-p)^{\lfloor k\rfloor}\sum_{i=1}^{\infty}p(1-p)^{i-1}\\
&=1-(1-p)^{\lfloor k\rfloor}
\ea\]
\epr
\item Mean:
\[E(X)=\frac{1}{p}\]
\bpr
\[\ba
E(X)&=\sum_{k=1}^{\infty}kp(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}(k-1)p(1-p)^{k-1}+\sum_{k=1}^{\infty}p(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}kp(1-p)^k+\sum_{k=1}^{\infty}p(1-p)^{k-1}
\ea\]
\[(1-p)E(X)=\sum_{k=1}^{\infty}kp(1-p)^k\]
\[E(X)-(1-p)E(X)=pE(X)=\sum_{k=1}^{\infty}p(1-p)^{k-1}=\frac{p}{1-(1-p)}=1\]
\epr
\item Variance:
\[\operatorname{Var}(X)=\frac{1-p}{p^2}\]
\bpr
\[\ba
E(X^2)&=\sum_{k=1}^{\infty}k^2p(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}(k-1)^2p(1-p)^{k-1}+2\sum_{k=1}^{\infty}kp(1-p)^{k-1}-\sum_{k=1}^{\infty}p(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}k^2p(1-p)^k+2E(X)-1
\ea\]
\[(1-p)E(X^2)=\sum_{k=1}^{\infty}k^2p(1-p)^k\]
\[E(X^2)-2E(X)+1-(1-p)E(X^2)=pE(X^2)-2E(X)+1=0\]
\[E(X^2)=\frac{2E(X)-1}{p}=\frac{2-p}{p^2}\]
\[\operatorname{Var}(X)=E(X^2)-E(X)^2=\frac{1-p}{p^2}\]
\epr
\item MGF:
\[M_X(t)=\frac{pe^t}{1-(1-p)e^t},\quad t<-\ln(1-p)\]
\bpr
\[\ba
E(e^{tX})&=\sum_{k=1}^{\infty}e^{kt}p(1-p)^{k-1}\\
&=\frac{pe^{t}}{1-(1-p)e^t}
\ea\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=1\]
\bpr
\[p_X(k+1)-p_X(k)=p(1-p)^k-p(1-p)^{k-1}=-p^2(1-p)^{k-1}<0\]
\epr
\item Median:
\[\operatorname{Median}(X)=\bcs\qty[\frac{-\ln 2}{\ln(1-p)},\frac{-\ln 2}{\ln(1-p)}+1],\quad&\frac{-\ln 2}{\ln(1-p)}\in\bbN\\\left\lceil\frac{-\ln 2}{\ln(1-p)}\right\rceil,\quad&\text{otherwise}\ecs\]
\bpr
\[P(X\le k)\ge\frac{1}{2}\]
\[1-(1-p)^{\lfloor k\rfloor}\ge\frac{1}{2}\]
\[(1-p)^{\lfloor k\rfloor}\le\frac{1}{2}\]
\[\lfloor k\rfloor\ge\frac{-\ln 2}{\ln(1-p)}\]
\[P(X\ge k)\ge\frac{1}{2}\]
\[\sum_{i=\lceil k\rceil}^{\infty}p(1-p)^{i-1}=(1-p)^{\lceil k\rceil-1}\sum_{i=1}^{\infty}p(1-p)^{i-1}=(1-p)^{\lceil k\rceil-1}\ge\frac{1}{2}\]
\[\lceil k\rceil-1\le\frac{-\ln 2}{\ln(1-p)}\]
\epr
\item Entropy:
\[H(X)=\frac{-p\ln p-(1-p)\ln(1-p)}{p}\]
\bpr
\[\ba
H(X)&=-\sum_{k=1}^{\infty}p(1-p)^{k-1}\ln(p(1-p)^{k-1})\\
&=-\sum_{k=1}^{\infty}p(1-p)^{k-1}\qty(\ln p+(k-1)\ln(1-p))\\
&=-\ln p-\ln(1-p)\sum_{k=1}^{\infty}kp(1-p)^{k-1}+\ln(1-p)\\
&=-\ln p-\ln(1-p)p\sum_{k=1}^{\infty}k(1-p)^{k-1}+\ln(1-p)
\ea\]
Let
\[\ba
S&=\sum_{k=1}^{\infty}k(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}(k-1)(1-p)^{k-1}+\sum_{k=1}^{\infty}(1-p)^{k-1}\\
&=\sum_{k=1}^{\infty}k(1-p)^k+\frac{1}{p}
\ea\]
\[(1-p)S=\sum_{k=1}^{\infty}k(1-p)^k\]
\[S-(1-p)S=pS=\frac{1}{p}\]
\[S=\frac{1}{p^2}\]
\[H(X)=-\ln p-\ln(1-p)p\frac{1}{p^2}+\ln(1-p)=\frac{-p\ln p-(1-p)\ln(1-p)}{p}\]
\epr
\item PMF monotonicity: strictly decreasing
\item Memorylessness:
\[P(X=j+k\mid X>j)=P(X=k)\quad j,k\in\bbN\]
\bpr
\[\frac{p(1-p)^{j+k}}{(1-p)^j}=p(1-p)^k.\]
\epr
\eit
\sssc{Negative binomial or Pascal random variable}
\bit
\item Trial: A negative binomial or Pascal random variable counts the number of independent Bernoulli trials with success probability $p$ until the $r$th success.
\item Notation: $X\sim\operatorname{NB}(r,p)$, $X\sim\operatorname{NegBin}(r,p)$, or $X\sim\operatorname{Pascal}(r,p)$, read as $X$ is negative binomial or Pascal distributed with parameters $r$ and $p$
\item Parameters: $r\in\bbN$, $p\in(0,1]$
\item Support: $\bbN\cap[r,\infty)$
\item PMF:
\[p_X(k)=\binom{k-1}{r-1}p^r(1-p)^{k-r},\quad k\in\bbN\cap[r,\infty)\]
\item CDF:
\[F_X(k)=\bcs0,\quad & k<r\\\frac{p^r}{(r-1)!}\sum_{i=0}^{\lfloor k\rfloor-r}(i+r-1)_i(1-p)^i,\quad&k\ge r\ecs\]
\item Mean:
\[E(X)=\frac{r}{p}\]
\bpr
Consider a sequence of i.i.d. Bernoulli trial with success probability $p$ with $T_i$ defined as the random variable of the number of trials until the $i$th success for $i\in\bbN\cap[1,r]$. Define $X_1=T_1$, and for $i\in\bbN\cap[2,r]$, define $X_i=T_i-T_{i-1}$. Then
\[X=T_r=\sum_{i=1}^rX_i.\]
And the $X_i$s are i.i.d. geometric random variable with parameter $p$. Thus, the mean of $X$ is $r$ times that of the geometric random variable.
\epr
\item Variance:
\[\operatorname{Var}(X)=\frac{r(1-p)}{p^2}\]
\item MGF:
\[M_X(t)=\qty(\frac{pe^{t}}{1-(1-p)e^t})^r,\quad t<-\ln(1-p)\]
\item Mode:
\[\operatorname{Mode}(X)=\left\lfloor\frac{r-1}{p}\right\rfloor\]
\bpr
\[\ba
p_X(k+1)-p_X(k)&=\binom{k}{r-1}p^r(1-p)^{k-r+1}-\binom{k-1}{r-1}p^r(1-p)^{k-r}\\
&=\qty(k(1-p)-(k-r+1))\frac{(k-1)!}{(r-1)!(k-r+1)!}p^r(1-p)^{k-r}\\
&=\qty(-kp+r-1)\frac{(k-1)!}{(r-1)!(k-r+1)!}p^r(1-p)^{k-r}
\ea\]
\[p_X(k+1)>p_X(k)\iff-kp+r-1>0\iff k<\frac{r-1}{p}\]
\[p_X(k+1)<p_X(k)\iff-kp+r-1<0\iff k>\frac{r-1}{p}\]
\epr
\item PMF monotonicity: strictly increasing for $k\le\frac{r-1}{p}$ and strictly decreasing for $k\ge\frac{r-1}{p}$
\item Decomposition: sum of $r$ i.i.d. geometric random variable with parameter $p$
\eit
\sssc{Poisson random variable}
\bit
\item Trial: A Poisson random variable counts the number of events occurring in a fixed interval of time in a Poisson (point) process, in which events occur with a known constant mean rate $\lambda$ and independently of the time since the last event.
\item Notation: $X\sim\operatorname{Pois}(\lambda)$ or $X\sim\operatorname{Poisson}(\lambda)$, read as $X$ is Poisson distributed with parameter or rate $\lambda$
\item Parameters: $\lambda\in(0,\infty)$
\item Support: $\bbN_0$
\item PMF:
\[p_X(k)=\frac{\lambda^ke^{-\lambda}}{k!}\]
\bpr
Let $N(t)$ denotes the number of events occuring in $[0,t]$, and let $p_k(t)=P(N(t)=k)$.
\[\lim_{h\to0^+}p_0(h)=1\]
\[\lim_{h\to0^+}\frac{p_1(h)}{h}=\lambda\]
\[\lim_{h\to0^+}\frac{p_k(h)}{h}=0,\quad k\in\bbZ\cap[2,\infty)\]
\[\ba
p_k'(t)&=\lim_{h\to0}\frac{p_k(t+h)-p_k(t)}{h}\\
&=\frac{p_k(t)(1-\lambda h)+p_{k-1}(t)(\lambda h)-p_k(t)}{h}\\
&=-\lambda p_k(t)+\lambda p_{k-1}(t)
\ea\]
\[p_0'(t)=-\lambda p_0(t)\]
\[p_0(t)=e^{-\lambda t}\]
\[p_1'(t)=-\lambda p_1(t)+\lambda e^{-\lambda t}\]
\[\mu(t)=e^{\int\lambda\dd{t}}=e^{\lambda t}\]
\[e^{\lambda t}p_1(t)=\lambda t\]
\[p_1(t)=\lambda te^{-\lambda t}\]
Assume
\[p_k(t)=\frac{(\lambda t)^ke^{-\lambda t}}{k!}.\]
Prove by mathematical induction. It holds for $k=0$ and $1$. Suppose it holds for $k\in\bbZ\cap[0,n]$. For $k=n+1$,
\[p_{n+1}'(t)=-\lambda p_{n+1}(t)+\frac{\lambda(\lambda t)^ne^{-\lambda t}}{n!}\]
\[\mu(t)=e^{\int\lambda\dd{t}}=e^{\lambda t}\]
\[e^{\lambda t}p_{n+1}(t)=\int\frac{\lambda(\lambda t)^n}{n!}\dd{t}=\frac{(\lambda t)^{n+1}}{(n+1)!}\]
\[p_{n+1}(t)=\frac{(\lambda t)^{n+1}e^{-\lambda t}}{(n+1)!}\]
\epr
\item CDF:
\[F_X(k)=e^{-\lambda}\sum_{i=0}^{\lfloor k\rfloor}\frac{\lambda^i}{i!}\]
\item Mean:
\[E(X)=\lambda\]
\bpr
\[\ba
E(X)&=\sum_{k=0}^{\infty}k\frac{\lambda^ke^{-\lambda}}{k!}\\
&=\lambda e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k}{k!}\\
&=\lambda e^{-\lambda}e^{\lambda}\\
&=\lambda
\ea\]
\epr
\item Variance:
\[\operatorname{Var}(X)=\lambda\]
\bpr
\[\ba
E(X^2)&=\sum_{k=0}^{\infty}k^2\frac{\lambda^ke^{-\lambda}}{k!}\\
&=\lambda e^{-\lambda}\sum_{k=0}^{\infty}(k+1)\frac{\lambda^k}{k!}\\
&=\lambda e^{-\lambda}\sum_{k=1}^{\infty}\frac{\lambda^k}{(k-1)!}+\lambda\\
&=\lambda^2+\lambda
\ea\]
\[\operatorname{Var}(X)=E(X^2)-E(X)^2=\lambda\]
\epr
\item MGF:
\[M_X(t)=\exp(\lambda\qty(e^t-1))\]
\bpr
\[\ba
E(e^{tX})&=\sum_{k=0}^{\infty}e^{kt}\frac{\lambda^ke^{-\lambda}}{k!}\\
&=e^{-\lambda}\sum_{k=0}^{\infty}\frac{\qty(\lambda e^t)^k}{k!}\\
&=e^{-\lambda}\exp(\lambda e^t)\\
&=\exp(\lambda\qty(e^t-1))
\ea\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=\lceil\lambda\rceil-1,\lfloor\lambda\rfloor\]
\bpr
\[\ba
p_X(k+1)-p_X(k)&=\frac{\lambda^{k+1}e^{-\lambda}}{(k+1)!}-\frac{\lambda^ke^{-\lambda}}{k!}\\
&=\qty(\lambda-k-1)\frac{\lambda^ke^{-\lambda}}{(k+1)!}
\ea\]
\[p_X(k+1)>p_X(k)\iff k<\lambda-1\]
\[p_X(k+1)<p_X(k)\iff k>\lambda-1\]
\epr
\item Entropy:
\[H(X)=\lambda\qty(1-\ln\lambda)+e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k\ln(k!)}{k!}\]
\bpr
\[\ba
H(X)&=-\sum_{k=0}^{\infty}\frac{\lambda^ke^{-\lambda}}{k!}\ln(\frac{\lambda^ke^{-\lambda}}{k!})\\
&=-e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k}{k!}\qty(k\ln\lambda-\lambda-\ln(k!))\\
&=-e^{-\lambda}\ln\lambda\sum_{k=0}^{\infty}k\frac{\lambda^k}{k!}+e^{-\lambda}\lambda\sum_{k=0}^{\infty}\frac{\lambda^k}{k!}+e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k\ln(k!)}{k!}\\
&=-e^{-\lambda}\lambda\ln\lambda\sum_{k=0}^{\infty}\frac{\lambda^k}{k!}+\lambda+e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k\ln(k!)}{k!}\\
&=\lambda\qty(1-\ln\lambda)+e^{-\lambda}\sum_{k=0}^{\infty}\frac{\lambda^k\ln(k!)}{k!}
\ea\]
\epr
\item PMF monotonicity: strictly increasing for $k\le\lambda-1$ and strictly decreasing for $k\ge\lambda-1$
\item Addictivity: For any $n\in\bbN$, for any $X_i\sim\operatorname{Pois}(\lambda_i)$ for all $i\in\bbN\cap[1,n]$,
\[\sum_{i=1}^nX_i\sim\operatorname{Pois}\qty(\sum_{i=1}^n\lambda_i)\]
\bpr
Suppose $X\sim\operatorname{Pois}(\lambda)$ and $Y\sim\operatorname{Pois}(\rho)$.
\[\ba
P(X_i+Y_i=k)&=\sum_{x=0}^kp_X(x)p_Y(k-x)\\
&=\sum_{x=0}^k\frac{\lambda^xe^{-\lambda}}{x!}\frac{\rho^{k-x}e^{-\rho}}{(k-x)!}\\
&=\frac{e^{-\lambda-\rho}}{k!}\sum_{x=0}^k\binom{k}{x}\lambda^x\rho^{k-x}\\
&=\frac{(\lambda+\rho)^ke^{-\lambda-\rho}}{k!}
\ea\]
\epr
\item Thinning property: Given $X\sim\operatorname{Pois}(\lambda)$ for some $\lambda\in(0,\infty)$, a random variable $Y$ such that $Y\mid(X=k)\sim\operatorname{Bin}(k,p)$ for some $p\in(0,1]$ must satisfy $Y\sim\operatorname{Pois}(\lambda p)$.
\bpr
\[\ba
P(Y=m)&=\sum_{k=1}^{\infty}P(Y=m\mid X=k)P(X=k)\\
&=\sum_{k=m}^{\infty}\binom{k}{m}p^m(1-p)^{k-m}\frac{\lambda^ke^{-\lambda}}{k!}
&=\frac{(\lambda p)^me^{-\lambda}}{m!}\sum_{k=m}^{\infty}\frac{(\lambda(1-p))^{k-m}}{(k-m)!}\\
&=\frac{(\lambda p)^me^{-\lambda}}{m!}e^{\lambda(1-p)}\\
&=\frac{(\lambda p)^me^{-\lambda p}}{m!}
\ea\]
\epr
\eit
\sssc{Poisson limit of binomials}
For any fixed $\lambda\in(0,\infty)$, let $X\sim\operatorname{Pois}(\lambda)$ and $X_n\sim\operatorname{Bin}\qty(n,\frac{\lambda}{n})$ for any $n\in\bbN$. Then for any $k\in\bbN_0$,
\[\lim_{n\to\infty}P(X_n=k)=P(X=k).\]
\bpr
\[\ba
\lim_{n\to\infty}P(X_n=k)&=\lim_{n\to\infty}\binom{n}{k}\frac{\lambda^k}{n^k}\qty(1-\frac{\lambda}{n})^{n-k}\\
&=\lim_{n\to\infty}\frac{n^k}{k!}\frac{\lambda^k}{n^k}e^{-\lambda}\qty(1-\frac{\lambda}{n})^{-k}\\
&=\lim_{n\to\infty}\frac{\lambda^k}{k!}e^{-\lambda}\qty(1-\frac{\lambda}{n})^{-k}\\
&=\frac{\lambda^ke^{-\lambda}}{k!}
\ea\]
\epr
\ssc{Families of Continuous Random Variables}
\sssc{Continuous or rectangular uniform random variable}
\bit
\item Notation: $X\sim\calU_{[a,b]}$, $X\sim\operatorname{Unif}(a,b)$, or $X\sim\operatorname{Uniform}(a,b)$, read as $X$ is continuously or rectangularly uniform distributed with parameters $a$ and $b$
\item Parameters: $a,b\in\bbR\land a<b$
\item Support: $[a,b]$
\item PDF:
\[f_X(x)=\frac{1}{b-a},\quad x\in[a,b]\]
\item CDF:
\[F_X(x)=\bcs 0,\quad&x<a\\ \frac{x-a}{b-a},\quad&x\in[a,b]\\1,\quad&x>b\ecs\]
\item Mean:
\[E(X)=\frac{a+b}{2}\]
\item Variance:
\[\operatorname{Var}(X)=\frac{(b-a)^2}{12}\]
\bpr
\[\ba
E(X^2)&=\int_a^b\frac{x^2}{b-a}\dd{x}\\
&=\frac{1}{b-a}\evlp(\frac{x^3}{3})_a^b\\
&=\frac{a^2+b^2+ab}{3}
\ea\]
\[\ba
\operatorname{Var}(X)&=E(X^2)-E(X)^2\\
&=\frac{a^2+b^2+ab}{3}-\frac{a^2+b^2+2ab}{4}\\
&=\frac{4a^2+4b^2+4ab-3a^2-3b^2-6ab}{12}\\
&=\frac{(b-a)^2}{12}
\ea\]
\epr
\item MGF:
\[M_X(t)=\bcs\frac{e^{bt}-e^{at}}{(b-a)t},\quad&t\ne 0\\1,\quad&t=0\ecs\]
\bpr
\[\ba
E(e^{tX})&=\int_a^b\frac{e^{tx}}{b-a}\dd{x}\\
&=\bcs\frac{e^{bt}-e^{at}}{(b-a)t},\quad&t\ne 0\\1,\quad&t=0\ecs
\ea\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=[a,b]\]
\item Median:
\[\operatorname{Median}(X)=\frac{a+b}{2}\]
\item Entropy:
\[h(X)=\ln(b-a)\]
\bpr
\[\ba
h(X)&=-\int_a^b\frac{1}{b-a}\ln(\frac{1}{b-a})\dd{x}\\
&=\ln(b-a)
\ea\]
\epr
\item Transition:
\[X-c\sim\operatorname{Unif}(a-c,b-c),\quad c\in\bbR\]
\item Scaling:
\[cX\sim\operatorname{Unif}(ca,cb),\quad c\in\bbR_{>0}\]
\[cX\sim\operatorname{Unif}(cb,ca),\quad c\in\bbR_{<0}\]
\item Symmetry: about $\frac{b-a}{2}$
\item PDF monotonicity: constant
\eit
\sssc{Normal (常態/正態) or Gaussian (高斯) random variable}
\bit
\item Notation:
\bit
\item $X\sim\calN(\mu,\sigma)$, $X\sim N(\mu,\sigma)$, or $X\sim\operatorname{Gaussian}(\mu,\sigma)$, read as $X$ is normal or Gaussian distributed with mean $\mu$ and standard deviation $\sigma$
\item $Z\sim\calN(0,1)$, $Z\sim N(0,1)$, or $X\sim\operatorname{Gaussian}(0,1)$ read as $X$ is standard or unit normal or Gaussian distributed
\eit
\item Parameters: $\mu\in\bbR$ and $\sigma\in(0,\infty)$
\item Support: $\bbR$
\item PDF:
\[\ba
f_X(x)&=\frac{1}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\\
&=\frac{1}{\sigma}\varphi\left(\frac{x-\mu}{\sigma}\right),
\ea\]
where $\varphi(x)$ is the PDF of $Z$, that is,
\[\varphi(x)=\frac{1}{\sqrt{2\pi}}\exp(-\frac{x^2}{2})\]
\item CDF:
\[F_X(x)=\Phi\qty(\frac{x-\mu}{\sigma}),\]
where $\Phi(x)$ is the CDF of $Z$, that is,
\[\Phi(x)=\frac{1}{2}\qty(1+\operatorname{erf}\qty(\frac{x}{\sqrt{2}})),\]
where $\operatorname{erf}(x)$ is the (Gaussian) error function, that is,
\[\erf(x)=\frac{2}{\sqrt{\pi}}\int_0^xe^{-t^2}\dd{t}\]
\bpr
Lemma:
\[I=\int_0^{\infty}e^{-t^2}\dd{t}=\frac{\sqrt{\pi}}{2}\]
Proof of lemma:
\[\ba
I^2&=\qty(\int_0^{\infty}e^{-x^2}\dd{x})\qty(\int_0^{\infty}e^{-y^2}\dd{y})\\
&=\int_0^{\infty}\int_0^{\infty}e^{-x^2}e^{-y^2}\dd{x}\dd{y}\\
&=\int_0^{\pi/2}\int_0^{\infty}e^{-r^2}r\dd{r}\dd{\theta}\\
&=\frac{1}{2}\int_0^{\pi/2}\int_0^{\infty}e^{-u}\dd{u}\dd{\theta},\quad u=r^2\\
&=\frac{1}{2}\int_0^{\pi/2}1\dd{\theta}\\
&=\frac{\pi}{4}
\ea\]
\[F_X(x)=\int_{-\infty}^x\frac{1}{\sqrt{2\pi}\sigma}\exp(-\frac{(t-\mu)^2}{2\sigma^2})\dd{t}\]
\[u=\frac{t-\mu}{\sqrt{2}\sigma},\quad\dd{u}=\frac{1}{\sqrt{2}\sigma}\dd{t}\]
\[\ba
F_X(x)&=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\frac{x-\mu}{\sqrt{2}\sigma}}e^{-u^2}\dd{u}\\
&=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\frac{x-\mu}{\sqrt{2}\sigma}}e^{-u^2}\dd{u}\\
&=\frac{1}{\sqrt{\pi}}\qty(\frac{\sqrt{\pi}}{2}+\int_0^{\frac{x-\mu}{\sqrt{2}\sigma}}e^{-u^2}\dd{u})\\
&=\frac{1}{2}+\frac{1}{\sqrt{\pi}}\int_0^{\frac{x-\mu}{\sqrt{2}\sigma}}e^{-u^2}\dd{u}\\
&=\frac{1}{2}\qty(1+\operatorname{erf}\qty(\frac{x-\mu}{\sqrt{2}\sigma}))
\ea\]
\epr
\item Quantile:
\[\ba
Q(p)&=\mu+\sigma\Phi^{-1}(p)\\
&=\mu+\sqrt{2}\sigma\operatorname{erf}^{-1}(2p-1)
\ea\]
\item Mean:
\[E(X)=\mu\]
\bpr
\[E(X)=\int_{-\infty}^{\infty}\frac{x}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\dd{x}\]
\[u=\frac{x-\mu}{\sqrt{2}\sigma},\quad\dd{u}=\frac{1}{\sqrt{2}\sigma}\dd{x}\]
\[\ba
E(X)&=\int_{-\infty}^{\infty}\frac{x}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\dd{x}\\
&=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}\qty(\sqrt{2}\sigma u+\mu)e^{-u^2}\dd{u}\\
&=\mu
\ea\]
\epr
\item Variance:
\[\operatorname{Var}(X)=\sigma^2\]
\bpr
\[E(X^2)=\int_{-\infty}^{\infty}\frac{x^2}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\dd{x}\]
\[u=\frac{x-\mu}{\sqrt{2}\sigma},\quad\dd{u}=\frac{1}{\sqrt{2}\sigma}\dd{x}\]
\[\ba
E(X^2)&=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}\qty(\sqrt{2}\sigma u+\mu)^2e^{-u^2}\dd{u}\\
&=\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}2\sigma^2u^2e^{-u^2}\dd{u}+\mu^2\\
&=\frac{2\sigma^2}{\sqrt{\pi}}\int_{-\infty}^{\infty}u^2e^{-u^2}\dd{u}+\mu^2\\
&=\frac{2\sigma^2}{\sqrt{\pi}}\qty(\evlp(-\frac{1}{2}ue^{-u^2})_{-\infty}^{\infty}+\int_{-\infty}^{\infty}ue^{-u^2}\dd{u})+\mu^2\\
&=-\frac{2\sigma^2}{\sqrt{\pi}}\qty(\evlp(-\frac{1}{2}e^{-u^2})_{-\infty}^{\infty}+\int_{-\infty}^{\infty}\frac{1}{2}e^{-u^2}\dd{u})+\mu^2\\
&=\sigma^2+\mu^2
\ea\]
\[\operatorname{Var}(X)=E(X^2)-E(X)^2=\sigma^2+\mu^2-\mu^2=\sigma^2\]
\epr
\item MGF:
\[M_X(t)=\exp(\mu t+\frac{\sigma^2t^2}{2})\]
\bpr
\[\ba
E(e^{tX})&=\frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu)^2}{2\sigma^2}+tx)\dd{x}\\
&=\frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu-\sigma^2t)^2}{2\sigma^2}+\frac{\sigma^2t^2}{2}+\mu t)\dd{x}\\
&=\frac{\exp(\frac{\sigma^2t^2}{2}+\mu t)}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu-\sigma^2t)^2}{2\sigma^2})\dd{x}\\
&=\exp(\mu t+\frac{\sigma^2t^2}{2})
\ea\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=\mu\]
\bpr
\[\ba
f_X'(x)&=-\frac{1}{\sqrt{2\pi}\sigma}\frac{2(x-\mu)}{2\sigma^2}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\\
&=-\frac{(x-\mu)}{\sqrt{2\pi}\sigma^3}\exp(-\frac{(x-\mu)^2}{2\sigma^2})
\ea\]
\[f_X'(x)>0\iff x<\mu\]
\[f_X'(x)<0\iff x>\mu\]
\epr
\item Median:
\[\operatorname{Median}(X)=\mu\]
\item Entropy:
\[h(X)=\frac{1}{2}\ln(2\pi e\sigma^2)\]
\bpr
\[u=\frac{x-\mu}{\sqrt{2}\sigma},\quad\dd{u}=\frac{1}{\sqrt{2}\sigma}\dd{x}\]
\[\ba
h(X)&=-\int_{-\infty}^{\infty}\frac{1}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\ln(\frac{1}{\sqrt{2\pi}\sigma}\exp(-\frac{(x-\mu)^2}{2\sigma^2}))\dd{x}\\
&=-\frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\qty(\ln(\frac{1}{\sqrt{2\pi}\sigma})-\frac{(x-\mu)^2}{2\sigma^2})\dd{x}\\
&=\frac{\ln(\sqrt{2\pi}\sigma)}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\dd{x}+\frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty}\exp(-\frac{(x-\mu)^2}{2\sigma^2})\frac{(x-\mu)^2}{2\sigma^2}\dd{x}\\
&=\frac{\ln(2\pi\sigma^2)}{2\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^2}\dd{u}+\frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}u^2e^{-u^2}\dd{u}\\
&=\frac{\ln(2\pi\sigma^2)}{2}+\frac{1}{\sqrt{\pi}}\qty(\evlp(-\frac{1}{2}ue^{-u^2})_{-\infty}^{\infty}+\frac{1}{2}\int_{-\infty}^{\infty}e^{-u^2}\dd{x})\\
&=\frac{\ln(2\pi\sigma^2)}{2}+\frac{1}{2}\\
&=\frac{1}{2}\ln(2\pi e\sigma^2)
\ea\]
\epr
\item Standardization:
\[X=\mu+\sigma Z\]
\item Transition:
\[X-c\sim\calN(\mu-c,\sigma),\quad c\in\bbR\]
\item Scaling:
\[cX\sim\calN(c\mu,c\sigma),\quad c\in\bbR_{\neq 0}\]
\item Symmetry: about $\mu$
\item PDF monotonicity: strictly increasing for $x\le\mu$ and strictly decreasing for $x\ge\mu$
\eit
\sssc{Exponential random variable}
\bit
\item Trial: An exponential random variable models the distance between adjacent events in a Poisson process with rate $\lambda$.
\item Notation: $X\sim\operatorname{Exp}(\lambda)$ or $X\sim\operatorname{Exponential}(\lambda)$, read as $X$ is exponential distributed with parameter or rate $\lambda$
\item Parameters: $\lambda\in(0,\infty)$
\item Support: $[0,\infty)$
\item PDF:
\[f_X(x)=\lambda e^{-\lambda x},\quad x\in\bbR_{\ge 0}\]
\bpr
Let $N(x)$ denotes the number of events occuring in $[0,x]$. We already know that
\[P(N(x)=k)=\frac{(\lambda x)^ke^{-\lambda t}}{k!}.\]
\[X=\inf\{x>0\mid N(x)\ge 1\}\]
\[\{X>x\}\iff\{N(x)=0\}.\]
\[P(X>x)=P(N(x)=0)=e^{-\lambda x}\]
\[F_X(x)=P(X\le x)=1-e^{-\lambda x}\]
\[f_X(x)=\lambda e^{-\lambda x}\]
\epr
\item CDF:
\[F_X(x)=\bcs0,\quad&x<0\\1-e^{-\lambda x},\quad&x\ge0\ecs\]
\item Quantile:
\[Q(p)=-\frac{\ln(1-p)}{\lambda}\]
\item Mean:
\[E(X)=\frac{1}{\lambda}\]
\bpr
\[\ba
E(X)&=\int_0^{\infty}x\lambda e^{-\lambda x}\dd{x}\\
&=\evlp(-xe^{-\lambda x})_0^{\infty}+\int_0^{\infty}e^{-\lambda x}\dd{x}\\
&=\evlp(-\frac{1}{\lambda}e^{-\lambda x})_0^{\infty}\\
&=\frac{1}{\lambda}
\ea\]
\epr
\item Variance:
\[\operatorname{Var}(X)=\frac{1}{\lambda^2}\]
\bpr
\[\ba
E(X^2)&=\int_0^{\infty}x^2\lambda e^{-\lambda x}\dd{x}\\
&=\evlp(-x^2e^{-\lambda x})_0^{\infty}+\int_0^{\infty}2xe^{-\lambda x}\dd{x}\\
&=\frac{2}{\lambda^2}
\ea\]
\[\operatorname{Var}(X)=E(X^2)-E(X)^2=\frac{2}{\lambda^2}-\frac{1}{\lambda^2}=\frac{1}{\lambda^2}\]
\epr
\item MGF:
\[M_X(t)=\qty(1-\frac{t}{\lambda})^{-1},\quad t<\lambda\]
\bpr
\[\ba
E(e^{tX})&=\int_0^{\infty}\lambda e^{(t-\lambda)x}\dd{x}\\
&=\evlp(\frac{\lambda e^{(t-\lambda)x}}{t-\lambda})_0^{\infty}\\
&=\frac{\lambda}{\lambda-t}\\
&=\qty(1-\frac{t}{\lambda})^{-1}
\ea\]
\epr
\item Mode:
\[\operatorname{Mode}(X)=0\]
\bpr
\[f_X'(x)=-\lambda^2e^{-\lambda x}<0\]
\epr
\item Median:
\[\operatorname{Median}(X)=\frac{\ln 2}{\lambda}\]
\item Entropy:
\[h(X)=1-\ln\lambda\]
\bpr
\[\ba
h(X)&=-\lambda\int_0^{\infty}e^{-\lambda x}\qty(\ln(\lambda)-\lambda x)\dd{x}\\
&=\evlp(e^{-\lambda x}\ln(\lambda))_0^{\infty}+\lambda^2\int_0^{\infty}xe^{-\lambda x}\dd{x}\\
&=1-\ln\lambda
\ea\]
\epr
\item Scaling:
\[cX\sim\operatorname{Exp}\qty(\frac{\lambda}{c}),\quad c\in\bbR_{>0}\]
\item Memorylessness:
\[P(X>x+y|X>x)=P(X>y),\quad x,y\in\bbR_{\ge0}\]
\bpr
\[\frac{e^{-\lambda (x+y)}}{e^{-\lambda x}}=e^{-\lambda y}\]
\epr
\item PDF monotonicity: strictly decreasing
\eit
\sssc{Erlang random variable}
\bit
\item Trial/Decomposition: An random variable with shape $k$ and rate $\lambda$ is the sum of $k$ i.i.d. exponential variables with rate $\lambda$.
\item Notation: $X\sim\operatorname{Erlang}(k,\lambda)$, read as $X$ is Erlang distributed with shape $k$ and rate $\lambda$, or parameters $k$ and $\lambda$
\item Parameters: $k\in\bbN$ and $\lambda\in(0,\infty)$
\item Support: $[0,\infty)$
\item PDF:
\[f_X(x)=\frac{\lambda^kx^{k-1}e^{-\lambda x}}{(k-1)!},\quad x\in\bbR_{\ge0}\]
\bpr
Prove by mathematical induction. For $k=1$,
\[f_1(x)=\lambda e^{-\lambda x}\]
Assume for $k=n$,
\[f_n(x)=\frac{\lambda^nx^{n-1}e^{-\lambda x}}{(n-1)!}\]
For $k=n+1$,
\[\ba
f_{n+1}(x)&=\int_0^xf_n(t)f_1(x-t)\dd{t}\\
&=\int_0^x\frac{\lambda^nt^{n-1}e^{-\lambda t}}{(n-1)!}\lambda e^{-\lambda(x-t)}\dd{t}\\
&=\int_0^x\frac{\lambda^{n+1}t^{n-1}e^{-\lambda x}}{(n-1)!}\dd{t}\\
&=\frac{\lambda^{n+1}x^ne^{-\lambda x}}{n!}
\ea\]
\epr
\item CDF:
\[F_X(x)=\bcs0,\quad&x<0\\1-\sum_{n=0}^{k-1}\frac{(\lambda x)^ne^{-\lambda x}}{n!},\quad&x\ge0\ecs\]
\bpr
\[\ba
F_k(x)&=\int_0^x\frac{\lambda^kt^{k-1}e^{-\lambda t}}{(k-1)!}\dd{t}\\
&=\frac{\lambda^{k-1}}{(k-1)!}\qty(\evlp(-t^{k-1}e^{-\lambda t})_0^x+(k-1)\int_0^xt^{k-2}e^{-\lambda t}\dd{t}
)\\
&=-\frac{(\lambda x)^{k-1}e^{-\lambda x}}{(k-1)!}+\frac{\lambda^{k-1}}{(k-2)!}\int_0^xt^{k-2}e^{-\lambda t}\dd{t}\\
&=-\frac{(\lambda x)^{k-1}e^{-\lambda x}}{(k-1)!}+F_{k-1}(x)
\ea\]
\[F_1(x)=1-e^{-\lambda x}\]
\[F_k(x)=1-\sum_{n=0}^{k-1}\frac{(\lambda x)^ne^{-\lambda x}}{n!}\]
\epr
\item Mean:
\[E(X)=\frac{k}{\lambda}\]
\item Variance:
\[\operatorname{Var}(X)=\frac{k}{\lambda^2}\]
\item MGF:
\[M_X(t)=\qty(1-\frac{t}{\lambda})^{-k}\]
\item Mode:
\[\operatorname{Mode}(X)=\frac{k-1}{\lambda}\]
\bpr
\[f_X'(x)=\frac{(k-1-\lambda x)\lambda^kx^{k-2}e^{-\lambda x}}{(k-1)!}\]
\[f_X'(x)>0\iff x<\frac{k-1}{\lambda}\]
\[f_X'(x)<0\iff x>\frac{k-1}{\lambda}\]
\epr
\item Scaling:
\[cX\sim\operatorname{Exp}\qty(k,\frac{\lambda}{c}),\quad c\in\bbR_{>0}\]
\item PDF monotonicity: strictly increasing for $x\le\frac{k-1}{\lambda}$ and strictly decreasing for $x\ge\frac{k-1}{\lambda}$
\eit
\sssc{Geometric to exponential limit theorem}
For any fixed $\lambda\in(0,\infty)$, let $X_n\sim\operatorname{Geom}\qty(\frac{\lambda}{n})$ and $Y_n=\frac{X_n}{n}$. Then
\[\lim_{n\to\infty}Y_n=\operatorname{Exp}(\lambda).\]
\bpr
\[F_{Y_n}(x)=1-\qty(1-\frac{\lambda}{n})^{\lfloor nx\rfloor}\]
\[\lim_{n\to\infty}F_{Y_n}(x)=1-e^{-\lambda x}\]
\epr
\sssc{Negative binomial to Erlang limit theorem}
For any fixed $\lambda\in(0,\infty)$, let $X_n\sim\operatorname{NegBin}\qty(r,\frac{\lambda}{n})$ and $Y_n=\frac{X_n}{n}$. Then
\[\lim_{n\to\infty}Y_n=\operatorname{Erlang}(r,\lambda).\]
\bpr
PLACEHOLDER
\epr
\end{document}
