\documentclass[a4paper,12pt]{article}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{3}
\input{/usr/share/LaTeX-ToolKit/template.tex}
\begin{document}
\title{Boolean Algebra}
\author{沈威宇}
\date{\temtoday}
\titletocdoc
\sct{Boolean Algebra}
\subsection{Introduction}
Boolean algebra, also called switching algebra, was introduced by George Boole in his first book The Mathematical Analysis of Logic (1847), and set forth more fully in his An Investigation of the Laws of Thought (1854).
\ssc{Values}
True (1 or T) and false (0 or F).
\subsection{Logical Operators}
\sssc{NOT / Negation / Inversion / Complement}
\begin{itemize}
\item Symbol: $A'$, $\neg A$, $\mathord{\sim}A$, or $\ol{A}$
\item Definition: One input, one output. True if the input is false and false otherwise.
\eit 
\sssc{AND / Conjunction / Logical Multiplication}
\begin{itemize}
\item Symbol: $A \cdot B$, $A*B$, $AB$, or $A\land B$
\item Definition: At least two inputs, one output. True if all inputs are true and false otherwise.
\eit
\sssc{OR / Disjunction / Logical Addition / Inclusive OR}
\begin{itemize}
\item Symbol: $A + B$ or $A\lor B$
\item Definition: At least two inputs, one output. True if at least one input is true and false otherwise.
\eit
\sssc{NAND / Not AND}
\begin{itemize}
\item Symbol: $A\uparrow B$
\item Definition: At least two inputs, one output. True if all inputs are false and false otherwise.
\eit
\sssc{NOR / Not OR}
\begin{itemize}
\item Symbol: $A\downarrow B$
\item Definition: At least two inputs, one output. True if at least one input is false and false otherwise.
\eit
\sssc{XOR / Exclusive OR}
\begin{itemize}
\item Symbol: $A \oplus B$
\item Definition: At least two inputs, one output. True if odd number of inputs are true and false otherwise.
\eit
\sssc{XNOR / Logical equivalence / Exclusive NOR / Biconditional}
\begin{itemize}
\item Symbol: $A\equiv B$, $A\odot B$, or $A\iff B$
\item Definition: At least two inputs, one output. True if even number (including 0) of inputs are true and false otherwise.
\eit
\sssc{IMPLY / Logical conditional / Material implication}
\begin{itemize}
\item Symbol: $A \implies B$
\item Definition: Two inputs, one output. $A\implies B$ is defined as $A'+B$.
\item Remark: If $A \implies B$, then $A$ is called a sufficient condition of $B$ or stronger than $B$, and $B$ is called a necessary condition of $A$ or weaket than $A$.
\eit
\sssc{NIMPLY / Material nonimplication}
\begin{itemize}
\item Symbol: $A\nimplies B$
\item Definition: Two inputs, one output. $A\nimplies B$ is defined as $AB'$.
\eit
\ssc{Boolean Arithmetic}
\sssc{Boolean expression}
A Boolean expression is formed by application of the logic operations to one or more Boolean variables or constants. 

\bit
\item\tb{Literals}: The simplest expressions, consisting of a single constant or variable or its complement, such as $0$, $X$, or $Y′$ .
\item\tb{Product terms}: A product term is a product of literals or a literal.
\item\tb{Sum terms}: A sum term is a sum of literals or a literal.
\eit

In a Boolean expression, parentheses are added as needed to specify the order in which the operations are performed; without parentheses, complementation is performed first, and then AND, and then OR; and the precedence of XOR and XNOR are either right higher to OR or right lower to OR depending on the context.
\sssc{Boolean function}
A Boolean function or a switching function is a function of which the domain is $\{1,0\}^n$ in which $n\in\mathbb{N}$ and the codomain is $\{1,0\}$.

The ON-set of a Boolean function $f$ is the set of all input combinations such that $f$ of them is $1$. Each element in the ON-set of $f$ corresponds to a row in the truth table of $f$ in which the output is $1$. The OFF-set of a Boolean function $f$ is the set of all input combinations such that $f$ of them is $0$. Each element in the OFF-set of $f$ corresponds to a row in the truth table of $f$ in which the output is $0$.
\sssc{Incompletely specified function (ISF) or Don't-care function}
An incompletely specified function (ISF) or a don't-care function is a Boolean function but in which for some input combinations, the output is not defined or irrelevant, often denoted as $X$. Those input combinations are called don't-care conditions.
\sssc{Truth table}
A truth table, also called a table of combinations, specifies the corresponding outputs for all possible input combinations of a Boolean function in order of the unsigned binary number corresponding to the input combination from small to large from top to bottom. A truth table for an $n$-variable Boolean function $f\qty(A_1,A_2,\ldots A_n)$ have $2^n$ rows and is:
\begin{longtable}[c]{|m|m|m|m|m|}
\hline
A_1 & A_2 & \ldots & A_n & f\\\hline
0 & 0 & \ldots & 0 & f\qty(0,0,\ldots 0)\\\hline
0 & 0 & \ldots & 1 & f\qty(0,0,\ldots 1)\\\hline
\vdots & \vdots & \vdots & \ddots & \vdots\\\hline
1 & 1 & \ldots & 1 & f\qty(1,1,\ldots 1)\\\hline
\end{longtable}
\sssc{Sum-of-products (SOP) form}
A Boolean expression is said to be in SOP form if it consists of a sum (OR) of product (AND) or single variable terms, in which it is called to be degenerate when some of the terms are single variables.

Every Boolean expression can be expressed in SOP form. And a Boolean expression may have more than one SOP forms.
\sssc{Product-of-sums (POS) form}
A Boolean expression is said to be in POS form if it consists of a product (AND) of sum (OR) or single variable terms, in which it is called to be degenerate when some of the terms are single variables.

Every Boolean expression can be expressed in POS form. And a Boolean expression may have more than one POS forms.
\sssc{Dual}
The dual of an Boolean expression or equation is another Boolean expression or equation obtained by simplifing the original expression or equation such that it only consists of literals, AND, and OR, and then replacing all AND in it with OR, all OR in it with AND, and all constants in it with their complements.
\sssc{Functionally complete}
A functionally complete set is a subset of the set of logic operators, 0, and 1 that can express all possible Boolean functions of any number of input variables. A minimal functionally complete set is a functionally complete set such that removing any one element from it makes it no longer functionally complete. Some minimal functionally complete sets are $\{\tx{AND},\tx{NOT}\}$, $\{\tx{OR},\tx{NOT}\}$, $\{\tx{NAND}\}$, $\{\tx{NOR}\}$, $\{\tx{IMPLY},\tx{0}\}$, and $\{\tx{NIMPLY},\tx{1}\}$, in which NAND and NOR are called universal.
\sssc{Boolean Vector Functions or Boolean Multi-output Functions}
A Boolean vector function or a Boolean multi-output function is a function $f\colon\{1,0\}^n\to\{1,0\}^m$. A Boolean vector function can be specified with truth table or as a tuple of $m$ Boolean functions.
\sssc{True and false as finite field or Galois field of two elements}
$(\{1,0\},\oplus,\land)$ where $\oplus$ denotes XOR and $\land$ denotes AND is finite field (aka Galois field) of two elements GF(2), also denoted as $\mathbb{F}_2$ or $\bbZ/2\bbZ$.
PLACEHOLDER
(after learned finite field (aka Galois field) in abstract algebra)
\sssc{Linear and Affine Boolean function}
A Boolean function $f\colon\{1,0\}^n\to\{1,0\}$ is linear iff
\[\forall a,b\in\{1,0\}^n\colon f(a\oplus b)=f(a)\oplus f(b),\]
where $\oplus$ denotes bitwise XOR, that is,
\[\exists a_1,a_2,\ldots,a_n\in\{1,0\}\colon f(x_1,x_2,\ldots,x_n)=\bigoplus_{i=1}^na_ix_i.\]

A Boolean function $f\colon\{1,0\}^n\to\{1,0\}$ is affine iff
\[\forall a,b\in\{1,0\}^n\colon f(a\oplus b)=f(a)\oplus f(b)\oplus f(0),\]
where $\oplus$ denotes bitwise XOR, that is,
\[\exists a_0,a_1,a_2,\ldots,a_n\in\{1,0\}\colon f(x_1,x_2,\ldots,x_n)=a_0\oplus\bigoplus_{i=1}^na_ix_i.\]

A Boolean vector function $f\colon\{1,0\}^n\to\{1,0\}^m$ is linear iff
\[\forall a,b\in\{1,0\}^n\colon f(a\oplus b)=f(a)\oplus f(b),\]
where $\oplus$ denotes bitwise XOR, that is, let the $k$th output of $f$ be a Boolean function $g_k$ of the inputs of $f$, for all $1\leq k\leq m$, $g_k$ is a linear Boolean function.

A Boolean vector function $f\colon\{1,0\}^n\to\{1,0\}^m$ is affine iff
\[\forall a,b\in\{1,0\}^n\colon f(a\oplus b)=f(a)\oplus f(b)\oplus f(0),\]
where $\oplus$ denotes bitwise XOR, that is, let the $k$th output of $f$ be a Boolean function $g_k$ of the inputs of $f$, for all $1\leq k\leq m$, $g_k$ is an affine Boolean function.

Some sources, however, use the term linear to refer to affine.
\sssc{Algebraic normal form (ANF), ring sum normal form (RSNF or RNF), Zhegalkin normal form, or Reed–Muller expansion}
A Boolean expression is said to be in algebraic norma form if it consists of a XOR of non-invertered AND or single variable terms and optionally a constant 1 term.

Every Boolean expression can be expressed in ANF. And a Boolean expression may have more than one POS forms.
\ssc{Theorems}
\sssc{Idempotent Laws}
\[XX = X,\quad X + X = X.\]
\sssc{Involution Law}
\[(X′)′ = X.\]
\sssc{Laws of Complementarity}
\[X X′ = 0,\quad X + X′ = 1.\]
\sssc{Commutativity of AND, OR, XOR, and XNOR}
\[X Y=Y X,\quad X+Y=Y+X,\]
\[X\oplus Y=Y\oplus X,\quad X\odot Y=Y\odot X.\]
\sssc{Associativity of AND, OR, XOR, and XNOR}
\[X Y Z=X (Y Z),\quad X+Y+Z=X+(Y+Z),\]
\[X\oplus Y\oplus Z=X\oplus (Y\oplus Z),\quad X\odot Y\odot Z=X\odot (Y\odot Z).\]
\sssc{Distributivity of AND over OR and OR over AND}
\[X (Y+Z)=X Y+X Z.\]
\[X+YZ=(X+Y)(X+Z).\]
\sssc{XOR of ANDs Theorem or Distributivity of AND over XOR}
\[X Y\oplus X Z=X (Y\oplus Z).\]
\begin{proof}
\[X Y\oplus X Z=XY(XZ)'+(XY)'XZ=XYX'+XYZ'+X'XZ+Y'XZ=XYZ'+XY'Z=X(Y\oplus Z).\]
\end{proof}
\sssc{XOR of ORs Theorem}
\[(X+Y)\oplus (X+Z)=X' (Y\oplus Z).\]
\begin{proof}
\[(X+Y)\oplus (X+Z)=(X+Y)(X+Z)'+(X+Y)'(X+Z)=XX'Z'+YX'Z'+X'Y'X+X'Y'Z=X'YZ'+X'Y'Z=X'(Y\oplus Z).\]
\end{proof}
\sssc{XNOR of ORs Theorem or Distributivity of OR over XNOR}
\[(X+Y)\odot (X+Z)=X+(Y\odot Z).\]
\begin{proof}
\[(X+Y)\odot (X+Z)=(X+Y)(X+Z)+(X+Y)'(X+Z)'=X+YZ+X'Y'Z'=X+YZ+Y'Z'=X+(Y\odot Z).\]
\end{proof}
\sssc{XNOR of ANDs Theorem}
\[XY\odot XZ=X'+(Y\odot Z).\]
\begin{proof}
\[XY\odot XZ=XYZ+(XY)'(XZ)'=XYZ+(X'+Y')(X'+Z')=XYZ+X'+Y'Z'=X'+(Y\odot Z).\]
\end{proof}
\sssc{DeMorgan's Laws}
\[(\sum_{i=1}^nX_i)′=\prod_{i=1}^nX_i',\quad (\prod_{i=1}^nX_i)'=\sum_{i=1}^nX_i',\]
\[(\bigoplus_{i=1}^nX_i)′=\bigodot_{i=1}^nX_i',\quad (\bigodot_{i=1}^nX_i)'=\bigoplus_{i=1}^nX_i'.\]
\sssc{Axiom of Equality}
For an one-to-one Boolean function $f$, a Boolean expression $A$ equals another Boolean expression $B$ if and only if $f(A)$ equals $f(B)$.
\sssc{Duality Principle}
A Boolean equation is an identity if and only if the dual of it is an identity.
\sssc{Uniting Theorems}
\[XY+XY'=X,\quad (X+Y)(X+Y')=X.\]
\sssc{Absorption Theorems}
\[X+XY=X,\quad X(X+Y)=X.\]
\sssc{Elimination Theorems}
\[X+X'Y=X+Y,\quad X(X'+Y)=XY.\]
\sssc{Consensus Theorems}
The consensus theorems involve eliminate one term from an expression in SOP or POS form, in which the eliminated term is called the consensus term.
\[XY+X′Z+YZ=XY+X′Z.\]
\begin{proof}
\[\begin{aligned}
XY+X′Z+YZ&=XY+X'Z+(X+X')YZ\\
&=XY+X'Z+XYZ+X'YZ\\
&=XY+X′Z.
\end{aligned}\]
\end{proof}
\[(X+Y)(X′+Z)(Y+Z)=(X+Y)(X′+Z).\]
\begin{proof}
\[\begin{aligned}
(X+Y)(X′+Z)(Y+Z)&=(X+Y)(X'+Z)(X+X')(Y+Z)\\
&=(X+Y)(X'+Z)(X+Y+Z)(X'+Y+Z)\\
&=(X+Y)(X′+Z).
\end{aligned}\]
\end{proof}
\sssc{Shannon's expansion theorem, Shannon decomposition, Boole's expansion theorem, or fundamental theorem of Boolean algebra}
The theorem states that
\[F=x\cdot F_x+x'\cdot F_{x'},\]
where $F$ is any $n$-variable Boolean function, $x$ is any independent variable of $F$, $F_x$ and $F_{x'}$, sometimes called the positive and negative Shannon cofactors, respectively, of $F$ with respect to $x$, are $(n-1)$-variable Boolean functions defined as $F$ with the $x$ set to $1$ and to $0$ respectively, and the RHS is called a Shannon's expansion or Boole's expansion of $F$.
\sssc{Combination of Distributivity and Consensus Theorem}
\[(X+Y)(X'+Z)=XZ+X'Y\]
\begin{proof}
\[(X+Y)(X'+Z)=0+XZ+X'Y+YZ=XZ+X'Y\]
\end{proof}
\sssc{XOR and XNOR Series Theorems}
\[\bigoplus_{i=1}^nX_i=\qty(\sum_{i=1}^nX_i)\mod 2.\]
\[\bigodot_{i=1}^nX_i=1-\qty(\sum_{i=1}^nX_i)\mod 2=\qty(1+\sum_{i=1}^nX_i)\mod 2.\]
\[\qty(\bigoplus_{i=1}^nX_i)'=\qty(\bigoplus_{i=1}^{j-1}X_i)\oplus\qty(X_j)'\oplus\qty(\bigoplus_{i=j+1}^nX_i),\quad\forall j\leq n\land j\in\mathbb{N},\,\forall n\text{\ s.t.\ }\frac{n}{2}\in\mathbb{N}.\]
\begin{proof}\mbox{}\\
Case $n=2$:
\[\ba
(X_1\oplus X_2)'&=(X_1X_2'+X_1'X_2)'=(X_1X_2')'(X_1'X_2)'\\
&=(X_1'+X_2)(X_1+X_2')=(X_1'X_2'+X_1X_2)\\
&=X_1'\oplus X_2=X_1\oplus X_2'
\ea\]
Prove by mathematical induction. Assume it holds for $n=k$ and $n=2$. We want to prove that it holds for $n=k+2$.
\[\ba
\qty(\bigoplus_{i=1}^{k+2})'&=\qty(\bigoplus_{i=1}^kX_i)'\oplus X_{k+1}\oplus X_{k+2}\\
&=\qty(\bigoplus_{i=1}^{j-1}X_i)\oplus\qty(X_j)'\oplus\qty(\bigoplus_{i=j+1}^kX_i)\oplus X_{k+1}\oplus X_{k+2},\quad \forall j\leq k\land j\in\mathbb{N}
\ea\]
\[\ba
\qty(\bigoplus_{i=1}^{k+2})'&=X_1\oplus X_2\oplus\qty(\bigoplus_{i=3}^{k+2}X_i)'\\
&=X_1\oplus X_2\oplus\qty(\bigoplus_{i=3}^{j-1}X_i)\oplus\qty(X_j)'\oplus\qty(\bigoplus_{i=j+1}^{k+2}X_i),\quad \forall 3\leq j\leq k+2\land j\in\mathbb{N}
\ea\]
\end{proof}
\[\bigoplus_{i=1}^nX_i=\qty(\bigodot_{i=1}^nX_i)',\quad\forall n\text{\ s.t.\ }\frac{n}{2}\in\mathbb{N}.\]
\begin{proof}\mbox{}\\
By the definitions, it holds for case $n=2$.

Prove by mathematical induction. Assume it holds for $n=k$ and $n=2$. We want to prove that it holds for $n=k+2$.
\[\ba
\bigoplus_{i=1}^{k+2}X_i&=\qty(\bigoplus_{i=1}^kX_i)\oplus X_{k+1}\oplus X_{k+2}\\
&=\qty(\bigodot_{i=1}^kX_i)'\oplus X_{k+1}\oplus X_{k+2}\\
&=\qty(\bigodot_{i=1}^{k+1}X_i)\oplus X_{k+2}\\
&=\qty(\bigodot_{i=1}^{k+2}X_i)'
\ea\]
\end{proof}
\[\bigoplus_{i=1}^nX_i=\bigodot_{i=1}^nX_i,\quad\forall n\text{\ s.t.\ }\frac{n-1}{2}\in\mathbb{N}.\]
\begin{proof}\mbox{}\\
By
\[\bigoplus_{i=1}^nX_i=\qty(\bigodot_{i=1}^nX_i)',\quad\forall n\text{\ s.t.\ }\frac{n}{2}\in\mathbb{N}.\]
For $n$ such that $\frac{n-1}{2}\in\mathbb{N}$,
\[\ba
\bigoplus_{i=1}^nX_i&=\bigoplus_{i=1}^{n-1}X_i\oplus X_n\\
&=\bigoplus_{i=1}^{n-1}X_i\oplus X_n\\
&=\qty(\bigodot_{i=1}^{n-1}X_i)'\oplus X_n\\
&=\bigodot_{i=1}^nX_i
\ea\]
\end{proof}
\sssc{Consensus of XOR Theorem}
\[X\oplus Y+X\oplus Z+Y\oplus Z=X\oplus Y+X\oplus Z=X\oplus Y+Y\oplus Z=X\oplus Z+Y\oplus Z=XY'+X'Z+YZ'=X'Y+XZ'+Y'Z\]
\begin{proof}
\[X\oplus Y+X\oplus Z=XY'+X'Y+XZ'+X'Z=XY'+X'Y+XZ'+X'Z+YZ'+Y'Z=X\oplus Y+X\oplus Z+Y\oplus Z=XY'+X'Z+YZ'=X'Y+XZ'+Y'Z\]
\end{proof}
\ssc{Minterm and Maxterm Expansions, Canonical Expansions, or Standard Expansion}
\sssc{Minterm}
The $n$-variable minterm of index $i$, denoted as $m_i$ is a function that outputs 1 if the unsigned binary number corresponding to the input combination is $i$ and outputs 0 otherwise.
\sssc{Maxterm}
The $n$-variable maxterm of index $i$, denoted as $M_i$ is a function that outputs 0 if the unsigned binary number corresponding to the input combination is $i$ and outputs 1 otherwise.
\sssc{Minterm expansion, (canonical/standard) sum-of-product (SOP) form, minterm canonical form, or canonical disjunctive normal form (CDNF)}
Minterm expansion is an expression of a completely specified Boolean function as a sum of minterms. Let $S$ be the set of all unsigned binary numbers corresponding to some input combination in the ON-set of $f$. The minterm expansion, canonical SOP, or standard SOP of $f$ is
\[f=\sum_{i\in S}m_i,\]
also denoted as
\[f=\sum m(i\in S)\]
where $i$ is listed from small to large.

For a given completely specified Boolean function, there exists a unique minterm expansion of it.

For example, given a function $f(a,b,c)$ with truth table
\begin{longtable}[c]{|m|m|m|m|}
\hline
a & b & c & f\\\hline
0 & 0 & 0 & 0\\\hline
0 & 0 & 1 & 0\\\hline
0 & 1 & 0 & 1\\\hline
0 & 1 & 1 & 0\\\hline
1 & 0 & 0 & 0\\\hline
1 & 0 & 1 & 1\\\hline
1 & 1 & 0 & 1\\\hline
1 & 1 & 1 & 0\\\hline
\end{longtable}
, the minterm expansion of it is
\[f=m_2+m_5+m_6=\sum m(2,5,6).\]
\sssc{Maxterm expansion, (canonical/standard) product-of-sum (POS) form, maxterm canonical form, or canonical conjunctive normal form (CCNF)}
Maxterm expansion is an expression of a completely specified Boolean function as a product of maxterms. Let $T$ be the set of all unsigned binary numbers corresponding to some input combination in the OFF-set of $f$. The maxterm expansion, canonical POS, or standard POS of $f$ is
\[f=\prod_{i\in T}M_i,\]
also denoted as 
\[f=\prod M(i\in T)\]
where $i$ is listed from small to large.

For a given completely specified Boolean function, there exists a unique maxterm expansion of it.

For example, given a function $f(a,b,c)$ with truth table
\begin{longtable}[c]{|m|m|m|m|}
\hline
a & b & c & f\\\hline
0 & 0 & 0 & 0\\\hline
0 & 0 & 1 & 0\\\hline
0 & 1 & 0 & 1\\\hline
0 & 1 & 1 & 0\\\hline
1 & 0 & 0 & 0\\\hline
1 & 0 & 1 & 1\\\hline
1 & 1 & 0 & 1\\\hline
1 & 1 & 1 & 0\\\hline
\end{longtable}
, the maxterm expansion of it is
\[f=M_0M_1M_3M_4M_7=\prod M(0,1,3,4,7).\]
\sssc{Incompletely specified boolean functions}
The definition of ON-set and OFF-set of an ISF is the same as completely specified function; the Don't-Care set or DC-set of an ISF is the set of all don't-care conditions.

The output for the don't-care conditions can be assigned either $0$ or $1$ during simplification or realization.

An ISF $f$ with ON-set $S$, OFF-set $T$, and DC-set $D$ is sometimes written similar to minterm expansion as
\[f=\sum m(i\in S)+\sum d(i\in D),\]
and similar to maxterm expansion as
\[f=\prod M(i\in T)\cdot\prod D(i\in D),\]
where each $d_i$ is called a don't-care terms.
\sssc{NOT, AND, and OR of functions}
For completely or incompletely specified Boolean function $f$:
\[f=\sum m(i\in S)+\sum d(i\in D)\iff f'=\prod M(i\in S)\cdot\prod D(i\in D).\]
\[f=\prod M(i\in T)\cdot\prod D(i\in D)\iff f'=\sum m(i\in T)+\sum d(i\in D).\]
For completely or incompletely specified Boolean functions $f$ and $g$ with same number of variables:
\[f=\sum m(i\in S_f)+\sum d(i\in D_f)\land g=\sum m(i\in S_g)+\sum d(i\in D_g)\implies fg=\sum m(i\in S_f\cap S_g)+\sum d\qty(i\in\qty(S_f\cup D_f)\cap\qty(S_g\cup D_g)\setminus\qty(S_f\cap S_g)).\]
\[f=\sum m(i\in S_f)+\sum d(i\in D_f)\land g=\sum m(i\in S_g)+\sum d(i\in D_g)\implies f+g=\sum m(i\in S_f\cup S_g)+\sum d\qty(i\in\qty(S_f\cup D_f)\cup\qty(S_g\cup D_g)\setminus\qty(S_f\cup S_g)).\]
\[f=\prod M(i\in T_f)\cdot\prod D(i\in D_f)\land g=\prod M(i\in T_f)\cdot\prod D(i\in D_f)\implies fg=\prod M(i\in T_f\cup T_g)\cdot\prod D\qty(i\in\qty(T_f\cup D_f)\cup\qty(T_g\cup D_g)\setminus\qty(T_f\cup T_g)).\]
\[f=\prod M(i\in T_f)\cdot\prod D(i\in D_f)\land g=\prod M(i\in T_f)\cdot\prod D(i\in D_f)\implies f+g=\prod M(i\in T_f\cap T_g)\cdot\prod D\qty(i\in\qty(T_f\cup D_f)\cap\qty(T_g\cup D_g)\setminus\qty(T_f\cap T_g)).\]
\ssc{Minimum form}
\sssc{Implicant or 1-term}
Given a Boolean function $f$ of $n$ variables, a product term $P$ is an implicant (aka 1-term) of $f$ iff for every combination of values of the $n$ variables for which $P = 1$, $f$ is also equal to $1$.
\sssc{Implicate or 0-term}
Given a Boolean function $f$ of $n$ variables, a sum term $P$ is an implicate (aka 0-term) of $f$ iff for every combination of values of the $n$ variables for which $P = 0$, $f$ is also equal to $0$.
\sssc{Prime implicant (PI)}
A prime implicant of a function $f$ is an implicant of $f$ which is no longer an implicant of $f$ if any literal is deleted from it.
\sssc{Prime implicate}
A prime implicate of a function $f$ is an implicate of $f$ which is no longer an implicate of $f$ if any literal is deleted from it.
\sssc{Essential prime implicant (EPI)}
An essential prime implicant of a function $f$ is a prime implicant $P$ of $f$ such that there exists a minterm $m$ of $f$ such that $m$ implies $P$ and for any other prime implicant $Q$ of $f$, $m$ doesn't implies $Q$.
\sssc{Essential prime implicate}
An essential prime implicate of a function $f$ is a prime implicate $P$ of $f$ such that there exists a maxterm $M$ of $f$ such that $P$ implies $M$ and for any other prime implicate $Q$ of $f$, $Q$ doesn't implies $M$.
\sssc{Minimum/minimal SOP form or minimum sum (of prime implicants)}
A SOP form of a Boolean function is called minimum if it has the fewest number of terms out of all SOP forms of the function, and every product term in it can not have any variable in it be eliminated. For a given Boolean function, there may exist more than one minimum SOP forms of it.

A minimum SOP form of a Boolean function must consist of some of its prime implicants (but not necessarily all). If a SOP form contains implicants that are not prime implicants, it is not a minimum SOP form.

A minimum SOP form of a Boolean function must contain all of its essential prime implicants.
\sssc{Minimum/minimal POS form or minimum product (of prime implicates)}
A POS form of a Boolean function is called minimum if it has the fewest number of terms out of all POS forms of the function, and every sum term in it can not have any variable in it be eliminated. For a given Boolean function, there may exist more than one minimum POS forms of it.

A minimum POS form of a Boolean function must consist of some of its prime implicates (but not necessarily all). If a POS form contains implicates that are not prime implicates, it is not a minimum POS form.

A minimum POS form of a Boolean function must contain all of its essential prime implicates.

One can simplify a Boolean function to minimum POS form by:
\ben
\item taking dual of the expression,
\item simplifying it to minimum SOP form, and
\item taking dual of it.
\een
\ssc{Karnaugh Maps}
A Karnaugh Map, aka a K-map, is a grid that visualize the ways to simplify a completely or incompletely specified Boolean function to a minimum SOP form.
\sssc{Draw Karnaugh Maps of less than five variables}
If the ouput of the input combination of a cell is $1$, we write $1$ in that cell; if the ouput of the input combination of a cell is $0$, we can leave that cell empty or write $0$ in that cell; if the input combination of a cell is a don't care condition, we write $X$ in that cell.

The order 00, 01, 11, 10 is the gray code order, which is such that adjacent cells in the K-map differ by only one variable.

The notation of a cell is the input combination of that cell (thus a binary number), called binary notation, or the binary integer representated by the input combination of that cell converted to decimal, called decimal notation.

A K-map for a $2$-variable Boolean function $f(A,B)$ is
\begin{longtable}[c]{c|c|c|}
\multicolumn{1}{c}{\thead{\backslashbox{$B$}{$A$}}} & \multicolumn{1}{c}{\thead{0}} & \multicolumn{1}{c}{\thead{1}} \\\cline{2-3}
\multicolumn{1}{c|}{\thead{0}} & $f(0,0)$ & $f(1,0)$ \\\cline{2-3}
\multicolumn{1}{c|}{\thead{1}} & $f(0,1)$ & $f(1,1)$ \\\cline{2-3}
\end{longtable}

A K-map for a $3$-variable Boolean function $f(A,B,C)$ is
\begin{longtable}[c]{c|c|c|}
\multicolumn{1}{c}{\thead{\backslashbox{$BC$}{$A$}}} & \multicolumn{1}{c}{\thead{0}} & \multicolumn{1}{c}{\thead{1}} \\\cline{2-3}
\multicolumn{1}{c|}{\thead{00}} & $f(0,0,0)$ & $f(1,0,0)$ \\\cline{2-3}
\multicolumn{1}{c|}{\thead{01}} & $f(0,0,1)$ & $f(1,0,1)$ \\\cline{2-3}
\multicolumn{1}{c|}{\thead{11}} & $f(0,1,1)$ & $f(1,1,1)$ \\\cline{2-3}
\multicolumn{1}{c|}{\thead{10}} & $f(0,1,0)$ & $f(1,1,0)$ \\\cline{2-3}
\end{longtable}

A K map for a $4$-variable Boolean function $f(A,B,C,D)$ is
\begin{longtable}[c]{c|c|c|c|c|}
\multicolumn{1}{c}{\thead{\backslashbox{$CD$}{$AB$}}} & \multicolumn{1}{c}{\thead{00}} & \multicolumn{1}{c}{\thead{01}} & \multicolumn{1}{c}{\thead{11}} & \multicolumn{1}{c}{\thead{10}} \\\cline{2-5}
\multicolumn{1}{c|}{\thead{00}} & $f(0,0,0,0)$ & $f(0,1,0,0)$ & $f(1,1,0,0)$ & $f(1,0,0,0)$ \\\cline{2-5}
\multicolumn{1}{c|}{\thead{01}} & $f(0,0,0,1)$ & $f(0,1,0,1)$ & $f(1,1,0,1)$ & $f(1,0,0,1)$ \\\cline{2-5}
\multicolumn{1}{c|}{\thead{11}} & $f(0,0,1,1)$ & $f(0,1,1,1)$ & $f(1,1,1,1)$ & $f(1,0,1,1)$ \\\cline{2-5}
\multicolumn{1}{c|}{\thead{10}} & $f(0,0,1,0)$ & $f(0,1,1,0)$ & $f(1,1,1,0)$ & $f(1,0,1,0)$ \\\cline{2-5}
\end{longtable}
\sssc{Group 1's}
The first step of simplifying a Boolean function using K-maps is grouping 1's (minterms).

A group (or loop) is a rectangle of $2^m\times 2^n$ cells that are either $1$ (minterm) or $X$ (don't-care condition), in which $m,n\in\mathbb{N}_0$, the leftmost and rightmost columns are horizontally adjacent, and the top and bottom rows are vertically adjacent. Cells can be in multiple groups.

Each group represents an implicant of $f$ by the following rules:
\ben
\item Let the set of all input variables that are $1$ in all cells in the group be $O$.
\item Let the set of all input variables that are $0$ in all cells in the group be $Z$.
\item The implicant represented by that group is $\prod_{o\in O}o\prod_{z\in Z}z'$.
\een
\sssc{Select collections of groups}
An allowed collection of groups must follow the following rules:
\bit
\item Any $1$ must be in at least one group.
\item If any $1$ is only covered by one possible group, then that group must be chosen, that is, any group representing an essential prime implicant must be chosen.
\item If any group $g$ is completely covered by another group, $g$ must not be chosen, that is, a group chosen must represent a prime implicant.
\eit
Find the allowed collections of groups that contain the fewest groups. For each one of them, the sum of the implicants corresponding to the groups in it is a minimum SOP form of the given function.
\sssc{Other uses of Karnaugh Maps}
K-maps have one-to-one correspondence to truth tables. We can perform logic operations on functions by perform the operation on each cell on their K-maps to obtain the K-map of the result function.
\sssc{Veitch Diagrams}
Veitch Diagrams are an alternative form of K-maps by labeling each input variable with a curly bracket containing the maximum rows or columns which that variable is 1 in the input combinations of all cells in them.
\sssc{Five-variable Karnaugh Maps}
One form of drawing a five-variable K-map is by placing one four-variable K-map on top of another, that is, for a $5$-variable Boolean function $f(A,B,C,D,E)$, the K-map is
{\fontsize{8pt}{12pt}\selectfont
\begin{longtable}[c]{cc|c|c|c|c|}
& \multicolumn{1}{c}{\thead{\backslashbox{$BC$}{$DE$}}} & \multicolumn{1}{c}{\thead{00}} & \multicolumn{1}{c}{\thead{01}} & \multicolumn{1}{c}{\thead{11}} & \multicolumn{1}{c}{\thead{10}} \\\cline{3-6}
& \multicolumn{1}{c|}{\thead{00}} & \backslashbox{$f(1,0,0,0,0)$}{$f(0,0,0,0,0)$} & \backslashbox{$f(1,0,1,0,0)$}{$f(0,0,1,0,0)$} & \backslashbox{$f(1,1,1,0,0)$}{$f(0,1,1,0,0)$} & \backslashbox{$f(1,1,0,0,0)$}{$f(0,1,0,0,0)$} \\\cline{3-6}
$A$ & \multicolumn{1}{c|}{\thead{01}} & \backslashbox{$f(1,0,0,0,1)$}{$f(0,0,0,0,1)$} & \backslashbox{$f(1,0,1,0,1)$}{$f(0,0,1,0,1)$} & \backslashbox{$f(1,1,1,0,1)$}{$f(0,1,1,0,1)$} & \backslashbox{$f(1,1,0,0,1)$}{$f(0,1,0,0,1)$} \\\cline{3-6}
\backslashbox{$1$}{$0$} & \multicolumn{1}{c|}{\thead{11}} & \backslashbox{$f(1,0,0,1,1)$}{$f(0,0,0,1,1)$} & \backslashbox{$f(1,0,1,1,1)$}{$f(0,0,1,1,1)$} & \backslashbox{$f(1,1,1,1,1)$}{$f(0,1,1,1,1)$} & \backslashbox{$f(1,1,0,1,1)$}{$f(0,1,0,1,1)$} \\\cline{3-6}
& \multicolumn{1}{c|}{\thead{10}} & \backslashbox{$f(1,0,0,1,0)$}{$f(0,0,0,1,0)$} & \backslashbox{$f(1,0,1,1,0)$}{$f(0,0,1,1,0)$} & \backslashbox{$f(1,1,1,1,0)$}{$f(0,1,1,1,0)$} & \backslashbox{$f(1,1,0,1,0)$}{$f(0,1,0,1,0)$} \\\cline{3-6}
\end{longtable}}
in which the adjacency between cells in the same layer of four-variable K-map is the same as that in four-variable K-map; the upper and lower triangles in the same square are adjacent; and a group is a cuboid of $2^l\times 2^m\times 2^n$ cells that are either $1$ (minterm) or $X$ (don't-care condition), in which $l,m,n\in\mathbb{N}_0$.

Another form is drawing the upper and lower triangles of the previous form as two four-variable K-maps side-by-side and drawing a line to connect two parts of a group on different map.

Yet another form is the same as the previous one but with Veitch diagram labeling and drawing two halves of boundary lines of a group surrounding two parts of a group on different maps instead of a line.
\sssc{Finding minimum POS form}
We can use K-maps to find minimum POS form of a completely or incompletely specified Boolean function by grouping $0$s instead of $1$s ($0$s and $X$s are allowed in a group), in which each group represents an implicate of $f$ by the following rules:
\ben
\item Let the set of all input variables that are $1$ in all cells in the group be $O$.
\item Let the set of all input variables that are $0$ in all cells in the group be $Z$.
\item The implicant represented by that group is $\sum_{o\in O}o'+\sum_{z\in Z}z$.
\een
, and the product of the sum terms corresponding to the groups in an allowed (covering all 0's) collections that contain the fewest groups is a minimum POS form of the Boolean function.
\ssc{Map-entered variable (MEV) or Variable entrant map (VEM)}
Map-entered variable (MEV), aka variable entrant map (VEM), is an extension of Karnaugh maps used to simplify certain kinds of completely or incompletely specified Boolean functions with typically more than four variables by allowing some variables to be entered as symbols or expressions inside the K-map cells rather than expanding the map's dimensions.
\sssc{One MEV}
Suppose we have an $n$-variable (usually five-variable) completely or incompletely specified Boolean function $f$ with $E$ being one of the input variable of $f$ being the MEV such that
\[\{f(P,1),f(P,0)\}\notin\{\{1,X\},\{0,X\}\},\]
for any $(n-1)$-variable input combination $P$, where $f(P,x)$ with $x\in\{1,0\}$ denotes the output of $f$ when the variables except MEVs be their corresponding values in $P$ and $E_i$ be $x$, that is, the function can be written as
\[f=\sum m(i\mid m_i\in A_1)+E\sum m(i\mid m_i\in A_E)+E'\sum m(i\mid m_i\in A_e)+\sum d(i\mid m_i\in A_X),\]
where $m_i$ is a possible $(n-1)$-variable minterm. Let $A$ be the set of all $(n-1)$-variable input combinations, and
\[A_0=A\setminus(A_1\cap A_E\cap A_e\cap A_X).\]
\ben
\item Draw a $(n-1)$-variable K-map for input variables except $E$. In each cell, let the $n-1$-variable input combination corresponding to the cell be $P$, where $f(P,E)$ denotes a function of $E$ where the variables except $E$ be their values in $P$:
\bit
\item If $f(P,1)=f(P,0)=1$, that is, the cell corresponds to a minterm in $A_1$, enter $1$.
\item If $f(P,1)=f(P,0)=0$, that is, the cell corresponds to a minterm in $A_0$, leave the cell empty or enter $0$.
\item If $f(P,1)=1$ and $f(P,0)=0$, that is, the cell corresponds to a minterm in $A_E$, enter $E$.
\item If $f(P,1)=0$ and $f(P,0)=1$, that is, the cell corresponds to a minterm in $A_e$, enter $E'$.
\item If $f(P,1)=f(P,0)=X$, that is, the cell corresponds to a minterm in $A_X$, enter $X$.
\eit
\item Group $1$s in same way as K-map ($1$s and $X$s are allowed in a group). Each group corresponding to a $(n-1)$-variable implicant.
\item Group $E$s in same way as K-map but treating $1$s as $X$s ($E$s, $1$s, and $X$s are allowed in a group). Each group corresponding to a $n$-variable implicant with $E$ be one of the literals.
\item Group $E'$s in same way as K-map but treating $1$s as $X$s ($E'$s, $1$s, and $X$s are allowed in a group). Each group corresponding to a $n$-variable implicant with $E'$ be one of the literals.
\item An allowed collection of groups must follow the following rules:
\bit
\item Any $1$, $E$, or $E'$ must be in at least one group.
\item If any $1$, $E$, or $E'$ is only covered by one possible group, then that group must be chosen, that is, any group representing an essential prime implicant must be chosen.
\item If any group $g$ is completely covered by another group, $g$ must not be chosen, that is, a group chosen must represent a prime implicant.
\eit
\item Find the allowed collections of groups that contain the fewest groups. For each one of them, the sum of the implicants corresponding to the groups in it is a minimum SOP form of the given function.
\een
\sssc{Multiple MEVs}
Suppose we have an $n$-variable completely or incompletely specified Boolean function $f$ in which there exist $m$ (usually $n-4$) variables $E_1,E_2,\ldots E_m$ chosen as MEVs such that
\[\{f(P,Q,1),f(P,Q,0)\}\notin\{\{1,X\},\{0,X\}\},\]
for any $(n-1)$-variable input combination $P$, for any fixed $(m-1)$-variable input combination $Q$, where $f(P,Q,x)$ with $x\in\{1,0\}$ denotes the output of $f$ when the variables except MEVs be their corresponding values in $P$, the MEVs except $E_i$ be their corresponding values in $Q$, and $E_i$ be $x$, that is, the function can be written as
\[f=\sum m(i\mid m_i\in A_1)+\sum_{i=1}^mE_i\sum m(i\mid m_i\in A_{E_i})+\sum_{i=1}^mE_i'\sum m(i\mid m_i\in A_{e_i})+\sum d(i\mid m_i\in A_X),\]
where $m_i$ is minterms in $(n-m)$-variable truth tables. Let $A$ be the set of all $(n-m)$-variable input combinations, and
\[A_0=A\setminus\qty(A_1\cap A_X\cap\bigcap_{i=1}^m\qty(A_{E_i}\cap A_{e_i})).\]
\ben
\item Draw a $(n-m)$-variable K-map for input variables except $E$ and $F$. In each cell, let the $(n-m)$-variable input combination corresponding to the cell be $P$, for any $i\in\mathbb{N}_{\leq m}$:
\bit
\item If $f(P,R)=1$ for any fixed $m$-variable input combination $R$ of the MEVs, where $f(P,R)$ denotes the output of $f$ when the variables except MEVs be their corresponding values in $P$ and the MEVs be their corresponding values in $R$, that is, the cell corresponds to a minterm in $A_1$, enter $1$.
\item If $f(P,R)=0$ for any fixed $m$-variable input combination $R$ of the MEVs, where $f(P,R)$ denotes the output of $f$ when the variables except MEVs be their corresponding values in $P$ and the MEVs be their corresponding values in $R$, that is, the cell corresponds to a minterm in $A_0$, enter $0$.
\item If the set of all possible values of $f(P,Q,1)$ is $\{1\}$ and the set of all possible values of $f(P,Q,0)$ is $\{0\}$, for any fixed $(m-1)$-variable input combination $Q$ of the MEVs except $E_i$, that is, the cell corresponds to a minterm in $A_{E_i}$, enter $E_i$.
\item If the set of all possible values of $f(P,Q,1)$ is $\{0\}$ and the set of all possible values of $f(P,Q,0)$ is $\{1\}$, for any fixed $(m-1)$-variable input combination $Q$ of the MEVs except $E_i$, that is, the cell corresponds to a minterm in $A_{e_i}$, enter $E_i'$.
\item If $f(P,R)=X$ for any fixed $m$-variable input combination $R$ of the MEVs, where $f(P,R)$ denotes the output of $f$ when the variables except MEVs be their corresponding values in $P$ and the MEVs be their corresponding values in $R$, that is, the cell corresponds to a minterm in $A_X$, enter $X$.
\eit
\item Group $1$s in same way as K-map ($1$s and $X$s are allowed in a group). Each group corresponding to a $(n-1)$-variable implicant.
\item For each $i\in\bbN_{\leq m}$, group $E_i$s in same way as K-map but treating $1$s as $X$s ($E_i$s, $1$s, and $X$s are allowed in a group). Each group corresponding to a $(n-m+1)$-variable implicant with $E_i$ be one of the literals.
\item For each $i\in\bbN_{\leq m}$, group $E_i'$s in same way as K-map but treating $1$s as $X$s ($E_i'$s, $1$s, and $X$s are allowed in a group). Each group corresponding to a $(n-m+1)$-variable implicant with $E_i'$ be one of the literals.
\item An allowed collection of groups must follow the following rules:
\bit
\item Any $1$, $E_i$, or $E_i'$ must be in at least one group.
\item If any $1$, $E_i$, or $E_i'$ is only covered by one possible group, then that group must be chosen, that is, any group representing an essential prime implicant must be chosen.
\item If any group $g$ is completely covered by another group, $g$ must not be chosen, that is, a group chosen must represent a prime implicant.
\eit
\item Find the allowed collections of groups that contain the fewest groups. For each one of them, the sum of the implicants corresponding to the groups in it is a minimum SOP form of the given function.
\een
\ssc{Quine-McCluskey Method}
The Quine-McCluskey method (along with the Petrick's method) is a systematic procedure to find the minimum SOP forms of a completely or incompletely specified Boolean function. Below, we first discuss about completely specified Boolean function.
\sssc{Minterm expansion}
First, we convert the function minterm expansion.

Below, we will take
\[f(a,b,c,d)=\sum m(0,1,2,5,6,7,8,9,10,14)\]
for example.
\sssc{Group minterms}
Group minterms with the same number of $1$'s in the input combination together and list them. Let the group of minterms with $k$ $1$'s be Group $k$. Group $k$ and Group $(k+1)$ are called adjacent.

For $f$:
\begin{longtable}[c]{|c|m|mmmm|}
\hline
Group 0 & 0 & 0 & 0 & 0 & 0 \\\hline
Group 1 & 1 & 0 & 0 & 0 & 1 \\\cline{2-6}
& 2 & 0 & 0 & 1 & 0 \\\cline{2-6}
& 8 & 1 & 0 & 0 & 0 \\\hline
Group 2 & 5 & 0 & 1 & 0 & 1 \\\cline{2-6}
& 6 & 0 & 1 & 1 & 0 \\\cline{2-6}
& 9 & 1 & 0 & 0 & 1 \\\cline{2-6}
& 10 & 1 & 0 & 1 & 0 \\\hline
Group 3 & 7 & 0 & 1 & 1 & 1 \\\cline{2-6}
& 14 & 1 & 1 & 1 & 0 \\\hline
\end{longtable}
\sssc{Combine minterms}
Compare minterms from adjacent groups and combine minterms that differ in only one input bit by
\[XY+XY'=X,\]
where $X$ is a product of literals and $Y$ is a literal.

(Note: It is obvious that two terms not in adjacent groups can't be combined by $XY+XY'=X$.)

Create next list by writing combined terms with $X$ preserved and $Y$, the differing bit, by $-$ (don't care), and writing not-touched terms as is.

Repeat combinations and create next list until no further combining is possible. The terms in this last list are the prime implicants of $f$.

For $f$:

List 1:
\begin{longtable}[c]{|c|m|mmmm|}
\hline
Group 0 & 0 & 0 & 0 & 0 & 0 \\\hline
Group 1 & 1 & 0 & 0 & 0 & 1 \\\cline{2-6}
& 2 & 0 & 0 & 1 & 0 \\\cline{2-6}
& 8 & 1 & 0 & 0 & 0 \\\hline
Group 2 & 5 & 0 & 1 & 0 & 1 \\\cline{2-6}
& 6 & 0 & 1 & 1 & 0 \\\cline{2-6}
& 9 & 1 & 0 & 0 & 1 \\\cline{2-6}
& 10 & 1 & 0 & 1 & 0 \\\hline
Group 3 & 7 & 0 & 1 & 1 & 1 \\\cline{2-6}
& 14 & 1 & 1 & 1 & 0 \\\hline
\end{longtable}

List 2:
\begin{longtable}[c]{|c|m|mmmm|}
\hline
Group 0 & 0,1 & 0 & 0 & 0 & - \\\cline{2-6}
& 0,2 & 0 & 0 & - & 0 \\\cline{2-6}
& 0,8 & - & 0 & 0 & 0 \\\hline
Group 1 & 1,5 & 0 & - & 0 & 1 \\\cline{2-6}
& 1,9 & - & 0 & 0 & 1 \\\cline{2-6}
& 2,6 & 0 & - & 1 & 0 \\\cline{2-6}
& 2,10 & - & 0 & 1 & 0 \\\cline{2-6}
& 8,9 & 1 & 0 & 0 & - \\\cline{2-6}
& 8,10 & 1 & 0 & - & 0 \\\hline
Group 2 & 5,7 & 0 & 1 & - & 1 \\\cline{2-6}
& 6,7 & 0 & 1 & 1 & - \\\cline{2-6}
& 6,14 & - & 1 & 1 & 0 \\\cline{2-6}
& 10,14 & 1 & - & 1 & 0 \\\hline
\end{longtable}

List 3:
\begin{longtable}[c]{|c|m|mmmm|}
\hline
Group 0 & 0,1,8,9 & - & 0 & 0 & - \\\cline{2-6}
& 0,2,8,10 & - & 0 & - & 0 \\\hline
Group 1 & 1,5 & 0 & - & 0 & 1 \\\cline{2-6}
& 2,6,10,14 & - & - & 1 & 0 \\\hline
Group 2 & 5,7 & 0 & 1 & - & 1 \\\cline{2-6}
& 6,7 & 0 & 1 & 1 & - \\\hline
\end{longtable}

Sometimes, all rows that have appeared in any of the previous lists are omitted and all rows that have been combined into any row in the next list are checked or crossed out. The terms that are not checked or crossed out are the prime implicants of $f$.

For $f$:

List 1:
\begin{longtable}[c]{|c|m|mmmm|c|}
\hline
Group 0 & 0 & 0 & 0 & 0 & 0 & \checkmark\\\hline
Group 1 & 1 & 0 & 0 & 0 & 1 & \checkmark\\\cline{2-6}
& 2 & 0 & 0 & 1 & 0 & \checkmark\\\cline{2-6}
& 8 & 1 & 0 & 0 & 0 & \checkmark\\\hline
Group 2 & 5 & 0 & 1 & 0 & 1 & \checkmark\\\cline{2-6}
& 6 & 0 & 1 & 1 & 0 & \checkmark\\\cline{2-6}
& 9 & 1 & 0 & 0 & 1 & \checkmark\\\cline{2-6}
& 10 & 1 & 0 & 1 & 0 & \checkmark\\\hline
Group 3 & 7 & 0 & 1 & 1 & 1 & \checkmark\\\cline{2-6}
& 14 & 1 & 1 & 1 & 0 & \checkmark\\\hline
\end{longtable}

List 2:
\begin{longtable}[c]{|c|m|mmmm|c|}
\hline
Group 0 & 0,1 & 0 & 0 & 0 & - & \checkmark\\\cline{2-6}
& 0,2 & 0 & 0 & - & 0 & \checkmark\\\cline{2-6}
& 0,8 & - & 0 & 0 & 0 & \checkmark\\\hline
Group 1 & 1,5 & 0 & - & 0 & 1 &\\\cline{2-6}
& 1,9 & - & 0 & 0 & 1 & \checkmark\\\cline{2-6}
& 2,6 & 0 & - & 1 & 0 & \checkmark\\\cline{2-6}
& 2,10 & - & 0 & 1 & 0 & \checkmark\\\cline{2-6}
& 8,9 & 1 & 0 & 0 & - & \checkmark\\\cline{2-6}
& 8,10 & 1 & 0 & - & 0 & \checkmark\\\hline
Group 2 & 5,7 & 0 & 1 & - & 1 &\\\cline{2-6}
& 6,7 & 0 & 1 & 1 & - &\\\cline{2-6}
& 6,14 & - & 1 & 1 & 0 & \checkmark\\\cline{2-6}
& 10,14 & 1 & - & 1 & 0 & \checkmark\\\hline
\end{longtable}

List 3:
\begin{longtable}[c]{|c|m|mmmm|}
\hline
Group 0 & 0,1,8,9 & - & 0 & 0 & - \\\cline{2-6}
& 0,2,8,10 & - & 0 & - & 0 \\\hline
Group 1 & 2,6,10,14 & - & - & 1 & 0 \\\hline
\end{longtable}
\sssc{Create prime implicant chart}
Create the prime implicant chart, a chart where the column are minterms and rows are prime implicants and each cell is marked $\times$ if the corresponding minterm is implied by the corresponding prime implicant.

For $f$:
\begin{longtable}[c]{m|mmmmmmmmmm}
\hline
& 0 & 1 & 2 & 5 & 6 & 7 & 8 & 9 & 10 & 14 \\\hline
(0,1,8,9) b'c' & \times & \times & & & & & \times & \times & & \\
(0,2,8,10) b'd' & \times & & \times & & & & \times & & \times & \\
(1,5) a'c'd & & \times & & \times & & & & & & \\
(2,6,10,14) cd' & & & \times & & \times & & & & \times & \times \\
(5,7) a'bd & & & & \times & & \times & & & & \\
(6,7) a'bc & & & & & \times & \times & & & & \\
\end{longtable}
\sssc{Select ssential prime implicants}
When a prime implicant is selected for inclusion in the minimum SOP form, the chart is reduced by deleting the corresponding row and the columns which correspond to the minterms covered by that prime implicant.

If a column has only one $\times$, the prime implicant of that row is an essential prime implicant.

Select all essential prime implicants.

For $f$: Select $(0,1,8,9) b'c'$ to cover $m_9$, and select $(2,6,10,14) cd'$ to cover $m_{14}$.
\sssc{Petrick's method}
Given the reduced chart, number all prime implicants as $P_1,P_2,\ldots$.

For $f$:
\[P_1=a'c'd,\quad P_2=a'bd,\quad P_3=a'bc.\]

For each remaining minterm $m_i$, construct a sum term of all prime implicants that covers it, and let $P$ be the product of them.

For $f$:
\[P=(P_1+P_2)(P_2+P_3).\]

Expand $P$ into canonical SOP form (with each $P_i$ considered a literal).

For $f$:
\[P=P_1P_2+P_2+P_2P_3.\]

Reduce $P$ by applying
\[X+XY=X\]
until no more such reduction is possible.

For $f$:
\[P=P_2.\]

Each term in reduced $P$ represents a solution, that is, a set of prime implicants which covers all minterms in the reduced chart. The sets with the least number of literals in it are the minimum SOP forms of $f$.

For $f$: $P_2$ is the only term in reduced $P$, thus the only minimum SOP form is 
\[f=P_2=b'c'+cd'+a'bd.\]
\sssc{Incompletely specified Boolean functions}
For incompletely specified Boolean functions, we do the following change to the process for completely specified Boolean functions:
\ben
\item Treat don't-care terms as minterms when grouping minterm.
\item Omit don't-care terms columns when creating the prime implicant chart.
\een
\end{document}
