\section{Multivariable Calculus}
\subsection{Notation and convention}
\begin{itemize}
\item $\mathbf{0}$ or $0$ refers to the zero element of a space $V$, that is, 
\[\forall\mathbf{v}\in V\colon\mathbf{v}+\mathbf{0}=\mathbf{v}.\]
\item Independent variable vector: $\mathbf{x}=(x_1,x_2,\dots,x_n)$. For two dimensions also $\mathbf{x}=(x,y)$. For three dimensions also $\mathbf{x}=(x,y,z)$.
\item Basis: $\mathbf{e}_i$ is the unit vector in the direction of $x_i$. For two or three dimensions also $\mb{i}=\mb{e}_i$ and $\mb{j}=\mb{e}_2$. For three dimensions also $\mathbf{k}=\mb{e}_3$.
\item Vector field: A function, often denoted with bold uppercase Latin letter, whose domain is a subset of $\bbR^n$ and codomain is $\bbR^m$, whose $i$th component is often denoted as the same character with subscript $i$, e.g., $F_i$ or $\mb{F}_i$ for vector field $\mb{F}$.
\item Scalar field: A function, often denoted with non-bold Latin or Greek letter, whose domain is a subset of $\bbR^n$ and codomain is $\bbR$.
\item $k$th total derivative function operator: $D^k$
\item Dot product (點積) operator: $\cdot$
\item Cross product (叉積) operator: $\times$
\item Gradient (梯度) operator: $\nabla$ or $\operatorname{grad}$
\item Jacobian matrix (雅可比矩陣): $\nabla$, $\mathbf{J}(\mathbf{F})$, $J(\mathbf{F})$, $\mb{J}_{\mb{F}}$, or $J_{\mathbf{F}}$ for operand vector field $\mb{F}$.
\item Gram matrix, Gramian matrix, or Gramian: $\mathbf{G}(\mathbf{F})$, $G(\mathbf{F})$, $\mb{G}_{\mb{F}}$, or $G_{\mathbf{F}}$ for operand vector field $\mb{F}$.
\item Divergence (散度) operator: $\nabla \cdot$ or $\operatorname{div}$
\item Curl (旋度) operator (three dimension): $\nabla\times$ or $\operatorname{curl}$
\item Directional derivative (方向導數) operator: $(\mb{F}\cdot\nabla)\mb{G}$ or $D_{\mb{F}}\mb{G}$ for directional derivative of tensor field $\mb{G}$ along/in the direction of a vector field $\mb{F}\colon\bbR^n\to\bbR^n$. For constant vector $\mb{F}$, $D_{\mb{F}}\mb{G}$ notation is typically used and $\mb{F}$ is typically a unit vector.
\item Hessian (黑塞) (matrix): 
$D^2f$, $\nabla\nabla$, $\nabla^2f$, $\mathbf{H}(f)$, $H(f)$, $\mb{H}_f$, $H_f$, or $\nabla\otimes\nabla f$ for operand scalar field $f$.
\item Laplace operator (拉普拉斯算子) or Laplacian: $\nabla\cdot\nabla$, $\nabla^2$, or $\Delta$. For clarity, we'll use $\nabla\cdot\nabla$ here.
\item Vector and covector convention: We use column vector convention of vectors and row vector convention of covectors here, and tensor convention follows. Note that a derivative function of a scalar field is a covector. We use the convention that the notation of partial derivative w.r.t. a column vector $v$ means the column vector of the partial derivatives w.r.t. each component of $v$.
\item Line/path/general integral operator: $\int$
\item Double/surface integral operator: $\iint$
\item Triple/volume integration operator: $\iiint$
\item Closed line integral operator: $\oint$
\item Closed surface integral operator: $\oiint$
\item Arc length element: $\dd{\ell}$
\item Vector arc length element: $\dd{\boldsymbol{\ell}}$, which is defined as unit tangent vector multiplied by $\dd{\ell}$.
\item Surface element: $\dd{S}$
\item Vector surface element: $\dd{\mb{S}}$, which is defined as unit normal vector in the orientation multiplied by $\dd{S}$ and is only defined for orientable surface equipped with an orientation. For a vector field $\mb{F}$, $\iint_S\mb{F}\cdot\dd{\mb{S}}$ is called flux of $f$ across $S$.
\item Positive orientation: The boundary curve $\ell$ of a compact surface $S$, $\dd{\boldsymbol{\ell}}$ transverses counterclockwise around $S$ unless otherwise specified.
\item Upward/Positive orientation: In $\bbR^2$, a surface is oriented upward unless otherwise specified.
\item Outward/Positive orientation: The boundary surface $S$ of a compact region of the same dimension of the space is oriented outward unless otherwise specified.
\end{itemize}
\subsection{Nabla operators}
\subsubsection{Gradient and Jacobian matrix}
The gradient of a tensor field $f$ is
\[\nabla f = \pdv{f}{(x_1,x_2,\ldots,x_n)} = \begin{bmatrix}\qty(\pdv{f}{x_1})^{\top} & \qty(\pdv{f}{x_2})^{\top} & \cdots  & \qty(\pdv{f}{x_n})^{\top}\end{bmatrix},\]
that is, the transpose of the first-order derivative of $f$. Some sources define gradient as the transpose of the definition here.

The gradient of a scalar field is a vector field. The gradient of a vector field is a bilinear form field (i.e., also a matrix) and is also called the Jacobian matrix of it. Some sources use the term gradient only for scalar fields.

The $(i,j)$th entry of the Jacobian matrix of a vector field $\mb{F}$ is
\[(J(\mb{F}))_{ij}=\pdv{F_i}{x_j}.\]
The determinant of Jacobian matrix is called Jacobian determinant or Jacobian.

The notation
\[\pdv{(F_1,\ldots,F_m)}{(x_1,\ldots,x_n)}\]
denotes the Jacobian matrix or sometimes Jacobian of $\mb{F}$.
\subsubsection{Directional derivative (方向導數)}
For a tensor field $\mb{G}\colon U\subseteq\bbR^n\to X$ and $\mb{F}\colon V\subseteq\bbR^n\to\bbR^n$, the directional derivative of $\mb{G}$ in the direction of $\mb{F}$ is a function $((\mb{F}\cdot\nabla)\mb{G})\colon W\subseteq\bbR^n\to X$ defined as
\[(\mb{F}\cdot\nabla)\mb{G}\coloneq\lim_{h\to 0}\frac{\mb{G}(\mb{x}+h\mb{F}(\mb{x}))-\mb{G}(\mb{x})}{h},\]
where $W$ is a subset of $U\cap V$ where the definition is defined.

For points where $\mb{G}$ is differentiable,
\[(\mb{F}\cdot\nabla)\mb{G}=\sum_{i=1}^nF_i\pdv{\mb{G}}{x_i}=\mb{F}(D\mb{A})=\mb{F}\cdot(\nabla\mb{G}).\]
\sssc{Gram matrix, Gramian matrix, or Gramian}
The Gram matrix, Gramian matrix, or Gramian $G(\mb{F})$ of a vector field $\mb{F}$ is defined as the transpose of the Jacobian matrix of $\mb{F}$ right multiplied by the Jacobian matrix of $\mb{F}$
\[G(\mb{F})=J(\mb{F})^{\top}J(\mb{F}),\]
that is,
\[G_{ij}=\pdv{\mb{F}}{x_i}\cdot\pdv{\mb{F}}{x_j}.\]
The determinant of Gram matrix is called Gram determinant or Gramian.
\subsubsection{Divergence (散度)}
For an $n$-dimensional vector field $\mb{F}$,
\[\nabla\cdot\mb{F}=\sum_{i=1}^n\pdv{F_i}{x_i},\]
which is a scalar field.
\subsubsection{Curl (旋度) (three dimension)}
For a three-dimensional vector field $\mb{F}(x,y,z)$ defined on a subset of $\bbR^3$,
\[\ba
\nabla\times\mathbf{F}=&\begin{bmatrix}
\mathbf{i} & \mathbf{j} & \mathbf{k} \\
\frac{\partial}{\partial x} & \frac{\partial}{\partial y} & \frac{\partial}{\partial z} \\
F_1 & F_2 & F_3
\end{bmatrix}\\
=&\qty(\pdv{F_3}{y}-\pdv{F_y}{z},\pdv{F_1}{z}-\pdv{F_3}{x},\pdv{F_2}{x}-\pdv{F_1}{y})
\ea\]
\subsubsection{Curl (two dimension)}
For a two-dimensional vector field $\mb{F}(x,y)$ defined on a subset of $\bbR^2$, $\pdv{F_2}{x}-\pdv{F_1}{y}$ is called curl, and $F_1,F_2$ are often denoted $P,Q$ respectively.
\subsubsection{Hessian (黑塞) (matrix)}
The Hessian matrix $H(f)$ of a scalar field $f$ is its second derivative, which is a tensor field of type $(0,2)$, represented as a matrix, that is,
\[(H(f))_{ij}=\frac{\partial^2f}{\partial x_i\partial x_j},\]
that is,
\[H(f)=J((\nabla f)).\]
\subsubsection{Laplace operator (拉普拉斯算子)}
The Laplace operator applied to a scalar field $f$ is
\[\nabla\cdot\nabla f = \nabla \cdot (\nabla f) = \tr(H(f))=\sum_{i=1}^n\frac{\partial^2 f}{\partial x_i^{\phantom{i}2}}.\]
\subsubsection{Poisson's equation (卜瓦松 or 帕松 or 泊松方程)}
An equation in the form
\[\nabla\cdot\nabla f = g(\mathbf{x})\]
for scalar fields $f,g$.
\subsubsection{Laplace's equation (拉普拉斯方程)}
An equation in the form
\[\nabla\cdot\nabla f = 0\]
for scalar field $f$.

A twice differentiable scalar field $f$ is called a harmonic function if $f$ satisfies Laplace's equation.
\ssc{Multi-index notation (多重指標記號)}
\sssc{Definition}
Suppose there are \( n \) variables \( x_1, x_2, \ldots, x_n \), then a multi-index $\alpha$ is a vector of \( n \) nonnegative integers $\alpha_1, \alpha_2,\ldots,\alpha_n$:
\[\alpha = (\alpha_1, \alpha_2,\ldots,\alpha_n).\]
\sssc{Norm or length}
The space is equipped with norm or length $|\alpha|$ or \( \|\alpha\| \);
\[|\alpha| = \alpha_1 + \alpha_2 + \ldots\alpha_n,\]
i.e., one-norm.
\sssc{Factorial}
The factorial \( \alpha! \) of it is defined as:
\[\alpha! = \alpha_1! \cdot \alpha_2! \cdot \ldots \alpha_n!.\]
\sssc{Exponent}
If \( \mathbf{x} = (x_1, x_2,\ldots, x_n) \), $\mathbf{x}^\alpha$ is defined as:
\[\mathbf{x}^\alpha = x_1^{\alpha_1} \cdot x_2^{\alpha_2} \ldots  x_n^{\alpha_n}.\]
\sssc{Order}
For two multi-indices $\alpha=(\alpha_1, \alpha_2,\ldots, \alpha_n)$ and $\beta=(\beta_1, \beta_2,\ldots,\beta_n)$, $\alpha\le\beta$ if $\alpha_i\le\beta_i$ for all $i\in\bbZ\cap[1,n]$.
\ssc{Higher-order derivatives}
\sssc{Higher-order mixed partial derivatives}
For a tensor field $f$ defined on a subset of $\bbR^n$ and a multi-index $\alpha=(\alpha_1, \alpha_2,\ldots, \alpha_n)$, $D^\alpha f$ or $\partial^\alpha f$ is defined as
\[D^\alpha f = \frac{\partial^{|\alpha|} f}{\partial x_1^{\pht{1}\alpha_1} \partial x_2^{\pht{2}\alpha_2} \ldots  \partial x_n^{\pht{n}\alpha_n}}.\]
\sssc{Higher-order derivative}
The $k$th order derivative of $\mathbf{F}:\,\mathbb{R}^n\to\mathbb{R}^m$, denoted as $\mathbf{F}^{(k)}(\mathbf{x})$ or $D^{k}\mathbf{F}(\mathbf{x})$, is a $(\mathbb{R}^n)^k\to\mathbb{R}$ function, where $(\mathbb{R}^n)^k$ is a Cartesian product of $k$ copies of $\mathbb{R}^n$ vector, that is,
\[D^{k}\mathbf{F}(\mathbf{x})=\sum_{|\alpha|=k} \left(D^\alpha \mathbf{F}(\mathbf{a})\right).\]
In particular, the first-order derivative of $\mathbf{F}$ is the gradient of it.
\subsubsection{Taylor series or Taylor expansion}
Let $f\colon D\subseteq\mathbb{R}^m\to\mathbb{R}$ be a function and $I\subseteq D$ be an open ball centered at $a$ such that $f\in C^{\infty}(I)$, then the Taylor series of $f$ at $a$ (or about $a$, near $a$, or centered at $a$) is
\[\sum_{\alpha\in\mathbb{N}_0^{\pht{0}m}}\frac{D^{\alpha}f(a)}{\alpha !}(x-a)^{\alpha},\]
of which convergence is not guaranteed, and convergence to $f$ is not guaranteed even if it's convergent.

Taylor series of at $0$ is called Maclaurin series or Maclaurin expansion.
\subsubsection{Taylor polynomial}
Let $f\colon D\subseteq\mathbb{R}^m\to\mathbb{R}$ be a function and $I\subseteq D$ be an open ball centered at $a$ where $f$ is $k$ times differentiable, then the $k$-th degree (or order) Taylor polynomial (or approximation) of $f$ at $a$ (or about $a$, near $a$, or centered at $a$) is
\[T_n(x)=\sum_{\alpha\in\mathbb{N}_0^{\pht{0}m}\land|\alpha|\leq k}\frac{D^{\alpha}f(a)}{\alpha !}(x-a)^{\alpha},\]
in which the first degree one is also called linear approximation, or linearization, or in three dimensions, tangent plane (approximation), and
\[R_n(x)=f(x)-T_n(x)\]
is called remainder.

Taylor polynomial of at $0$ is called Maclaurin polynomial.
\ssc{Graph}
\sssc{Level curve, level slice, or contour line}
For a real-valued function $f$ of two variables, the level curve of $f$ at level $k\in\bbR$ is defined as the set of $\mb{x}$ such that
\[f(\mb{x})=k.\]
\sssc{Contour map or contour plot}
Contour map or contour plot is the graphical representation of the level curves of a function of two variable, typical represented with a collection of level curves labelled with corresponding levels that are often of same difference relative to their adjacent curves, or with color or height ($z$-axis) showing function value at all points.
\sssc{Level surface}
For a real-valued function $f$ of three variables, the level surface of $f$ at level $k\in\bbR$ is defined as the set of $\mb{x}$ such that
\[f(\mb{x})=k.\]
\sssc{Gradient perpendicular to tangent of level}
For function $F(\mb{x})\colon D\subseteq\bbR^n\to\bbR$, suppose $\mb{r}(t)$ is any parameterized curve lies on $F(\mb{x})=k$, then
\[\nabla F\cdot\mb{r}'=0.\]
\sssc{Tangent flat}
For function $F(\mb{x})\colon D\subseteq\bbR^n\to\bbR$, for each point $\mb{x}_0\in D$, the tangent flat of $F$ at it is given by
\[\nabla F\cdot(\mb{x}-\mb{x}_0)=0.\]
\sssc{Normal line}
For function $F(\mb{x})\colon D\subseteq\bbR^n\to\bbR$, for each point $\mb{x}_0\in D$, the normal line of $F$ at it is given by
\[\mb{x}(t)=t\nabla F(\mb{x}_0)+\mb{x}_0,\quad t\in\bbR.\]
\subsection{Line integral (線積分) or Path integral (路徑積分)}
\subsubsection{Definition}
For a parameterized curve $\gamma(t)\colon[a,b]\to\bbR^n$ that is piecewise $C^1$, the line integral of a function $f$ defined on the image of $\gamma$ over $\gamma$ is defined as the Riemann–Stieltjes integral:
\[\int_{\gamma}f(\mb{x})\dd{\mb(x)}\coloneq\int_{t=a}^bf(\gamma(t))\dd{\gamma(t)},\]
In which $\gamma$ is called the integral path.

The integral does not depend on the parametrization of the curve as long as the curve starts at $\gamma(a)$ and ends at $\gamma(b)$ and is piecewise $C^1$.
\sssc{Conservative field (保守場)}
An $n$-dimensional vector field $\mb{v}$ defined on an open subset $U$ of $\bbR^n$ is said to be conservative if there exists a $C^1$ scalar field $\varphi$ on $U$ such that
\[\mb{v}=\nabla\varphi\]
on $U$.
\sssc{Path independence}
Line integrals of an $n$-dimensional vector field $\mb{v}$ defined on an open subset $U$ of $\bbR^n$ is said to be path-independent or independent of path on $U$ if for any two piecewise $C^1$ paths $C,D\subseteq U$ with same initial and terminal points,
\[\int_C\mb{v}\cdot\dd{\mb{x}}=\int_D\mb{v}\cdot\dd{\mb{x}}\]
\ssc{Multiple integral}
\sssc{Integral region}
The integral region of $n$-tuple integral is a measurable subset of $\bbR^n$, often called plane region when $n=2$ and solid region when $n=3$.
\sssc{Lebesgue integral}
Defined via product measure space. The $\dd{\mu}$ where $\mu$ is Lebesgue measure is often denoted as $\dd{A}$ or sometimes $\dd{S}$ for two dimensions and $\dd{V}$ for three dimensions or sometimes higher dimensions.
\sssc{Riemann integral}
For rectangle
\[R=\prod_{i=1}^n[x_{i0},x_{i1}],\]
and function $f$ bounded on $R$,
\[\int_Rf(\mb{x})\dd{V}\coloneq\lim_{(m_1,m_2,\ldots,m_n)\to\infty}\qty(\sum_{j_i=1}^{m_i})_{i=1}^nf\qty(\mb{x}_{(j_i)_{i=1}^n}*)\Delta V,\]
called multiple Riemann sum, where
\[\Delta x_i=\frac{\qty(x_{i1}-x_{i0})}{m_i},\]
\[\Delta V=\prod_{i=1}^n\Delta x_i,\]
and $\mb{x}_{(j_i)_{i=1}^n}*$ is an arbitrary chosen sample point in
\[\prod_{i=1}^n\qty[x_{i0}+\qty(j_i-1)\Delta x_i,x_{i0}+j_i\Delta x_i.\]
For arbitrary subset $E\subseteq R$ and function $g$ bounded on $E$, let $h$ be the trivial extension of $g_E$ to $R$,
\[\int_Eg(\mb{x})\dd{V}\coloneq\int_Rh(\mb{x})\dd{V}.\]
\sssc{Iterated integral}
An integral whose integrand is an integral.
\ssc{Type of region}
\sssc{Simple region}
A $x_i$ simple region $V$ in $\bbR^n$ is a region that can be written in the form
\[V=\left\{(x_1,\ldots,x_n)\colon (x_{\substack{j=1\\j\neq i}}^n)\in D\land\varphi_1(x_{\substack{j=1\\j\neq i}}^n)\le x_i\le\varphi_2(x_{\substack{j=1\\j\neq i}}^n)\right\}.\]
where $\varphi_1,\varphi_2$ are continuous functions.
\sssc{Type of region in two dimensions}
A Type I region is a subset of $\bbR^2$ of the form
\[D=\{(x,y)\mid a\le x\le b\land g_1(x)\le y\le g_2(x)\}\]
where $g_1,g_2$ are continuous functions defined on $[a,b]$.

A Type II region is a subset of $\bbR^2$ of the form
\[D=\{(x,y)\mid a\le y\le b\land g_1(y)\le x\le g_2(y)\}\]
where $g_1,g_2$ are continuous functions defined on $[a,b]$.
\sssc{Type of region in three dimensions}
A Type I region is a subset of $\bbR^3$ of the form
\[D=\{(x,y,z)\mid (x,y)\in E\land g_1(x,y)\le z\le g_2(x,y)\}\]
where $E$ is a plane $g_1,g_2$ are continuous functions defined on $E$.

A Type II region is a subset of $\bbR^3$ of the form
\[D=\{(x,y,z)\mid (y,z)\in E\land g_1(y,z)\le x\le g_2(y,z)\}\]
where $E$ is a plane $g_1,g_2$ are continuous functions defined on $E$.

A Type III region is a subset of $\bbR^3$ of the form
\[D=\{(x,y,z)\mid (x,z)\in E\land g_1(x,z)\le y\le g_2(x,z)\}\]
where $E$ is a plane $g_1,g_2$ are continuous functions defined on $E$.
\ssc{Manifold element in terms of Hausdorff measure}
\sssc{Differentiable parametric manifold}
An $m$-dimensional differentiable manifold $S$ in $\bbR^n$ with $1\le m<n$ parameterized as $\mb{r}(t_1,t_2,\ldots,t_m)$ on domain $D$ has $m$-dimensional Hausdorff measure
\[\int_D\sqrt{\det(G)}\dd{t_1}\dd{t_2}\ldots\dd{t_m}\]
and manifold element
\[\dd{S}=\sqrt{\det(G)}\dd{t_1}\dd{t_2}\ldots\dd{t_m},\]
where $\det(G)$ is the Gram determinant of $\mb{r}$.
\sssc{Differentiable parametric surface}
A differentiable surface $S$ in $\bbR^3$ parameterized as $\mb{r}(s,t)$ on domain $D$ has two-dimensional Hausdorff measure
\[\int_D\abs{\pdv{\mb{r}}{s}\time\pdv{\mb{r}}{t}}\dd{s}\dd{t}\]
and surface element
\[\dd{S}=\abs{\pdv{\mb{r}}{s}\time\pdv{\mb{r}}{t}}\dd{s}\dd{t}.\]
If $S$ is orientable and equipped with an orientation, the vector surface element is
\[\dd{\mb{S}}=\pm\pdv{\mb{r}}{s}\time\pdv{\mb{r}}{t}\dd{s}\dd{t},\]
where $\pm$ is such that $\pm\UnitVector(\pdv{\mb{r}}{s}\time\pdv{\mb{r}}{t})$ is in the orientation.
\sssc{Hypersurface as graph}
An $(n-1)$-dimensional differentiable hypersurface $S$ in $\bbR^n$ parameterized w.r.t. $n-1$ of the Cartesian coordinates, say, $x_1,x_2,\ldots,x_{n-1}$, as $x_n=f(x_1,x_2,\ldots,x_{n-1}),\quad (x_1,x_2,\ldots,x_{n-1})\in D$, where $f\in W^{1,1}(\bbR)$, has $(n-1)$-dimensional Hausdorff measure
\[\int_D\sqrt{1+\abs{\nabla f}^2}\dd{\mu(x_1,x_2,\ldots,x_{n-1})}\]
and Hausdorff measure element
\[\dd{S}=\sqrt{1+\abs{\nabla f}^2}\dd{\mu(x_1,x_2,\ldots,x_{n-1})}\]
