\sct{Function of Bounded Variation}
\ssc{Of real functions}
\subsubsection{Partition of an interval}
A partition $P(x, n)$ of a compact interval $[a,b]$ is a finite sequence of numbers of the form
\[P(x, n):=\{x_i\colon x_0=a\land x_n=b\land\forall 1\leq i<j\leq n\colon x_i<x_j\}_{i=0}^n.\]
Each $[x_i, x_{i+1}]$ is called a subinterval in the partition. The length of a closed interval $[c,d]$ is defined as $d-c$. The mesh or norm of a partition is defined as
\[\max_i\left(x_{i+1}-x_{i}\right)\]
for every integer $i\in [0,n-1]$.
\sssc{Total variation}
Let $f\colon X\subseteq\bbR\to\bbR$ be a function. The total variation of $f$ over interval $[a,b]\subseteq X$, denoted as $V_a^b(f)$, $V_{[a,b]}(f)$, $V(f,[a,b])$, or $V(f;[a,b])$, is
\[V_a^b(f)=\sup_{P(x,n)\in\mathcal{P}}\sum_{i=0}^{n-1}\abs{f\qty(x_{i+1})-f\qty(x_i)},\]
where $\mathcal{P}$ is the set of all partitions of $f$ over $[a,b]$.
\sssc{Function of bounded variation (BV)}
$f$ is called of bounded variation or called a (globally) BV function over $[a,b]$ iff $V_a^b(f)$ is finite.
\sssc{Function of locally bounded variation}
$f$ is called of locally bounded variation or called a locally BV function over an open interval $(a,b)$ iff $V_c^d(f)$ is finite for all $[c,d]\in (a,b)$.
\sssc{Space of functions of bounded variation}
The space of all (globally) BV functions over $[a,b]$ is denoted as $\operatorname{BV}([a,b])$.
Banach?
PLACEHOLDE 
\sssc{Space of functions of bounded variation}
The space of all locally BV functions over $(a,b)$ is denoted as $\operatorname{BV}_{\tx{loc}}((a,b))$
Banach?
PLACEHOLDER
\sssc{Induced measure}
For a BV function $g\colon(a,b)\subseteq\bbR\to\bbR$, the signed measure $\mu_g(x)$ induced by $g$, is defined for any interval $(c,d)\subseteq (a,b)$, as
\[\mu_g((c,d))=g(d)-g(c).\]
If $g$ is non-decreasing, then $\mu_g$ is non-negative. If $g$ is non-increasing, then $\mu_g$ is non-positive.
\sssc{Jordan decomposition of BV function}
The Jordan decomposition $g^+$ and $g^-$ of a function $g$ of bounded variation on interval $[a,b]$ is defined as
\[g^+(x)=V_a^x(g),\quad g^-(x)=g(x)-g^+(x),\]
and are both non-decreasing.
\ssc{Of real fields}
PLACEHOLDER
\ssc{In measure theory}
PLACEHOLDER
